---
title: "R Notebook"
output: html_notebook
---

Proteomics analysis of FetuinB exprssing cells (WT vs NAFLD)

```{r}
#lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE)

library(data.table)
library(stringr)
library(openxlsx)
library(org.Hs.eg.db)
library(AnnotationDbi)
library(clusterProfiler)
library(ReactomePA)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(factoextra)
library(ggfortify)
library(colorspace)
library(RColorBrewer)
library(ggrepel)
library(EnhancedVolcano)
library(plyr)
```

###################################################### 
LOAD DATA
######################################################

Caressa's data
```{r}
protLFQ <- readxl::read_excel("Input/Summary_Results_HEP.xlsx", sheet="Perseus")
protLFQ_trim <- data.table(protLFQ)[, c("Gene names","Welch's T-test Difference LFQ intensity TRAb_LFQ intensity WTAb", "Welch's T-test q-value LFQ intensity TRAb_LFQ intensity WTAb")] %>% setNames(c("geneName", "LFQ_FetB.vs.WT", "qvalue"))
#note: can't include protein names because number of proteins names does NOT equal number of gene names for split
protLFQ_parse <- mutate(protLFQ_trim, id=as.character(1:nrow(protLFQ_trim))) %>% tidyr::separate_rows(geneName, sep=";") %>% data.table()

protLFQ_sig <- protLFQ_parse[qvalue <= 0.1 & abs(LFQ_FetB.vs.WT) >= log2(1.5) ,][order(LFQ_FetB.vs.WT)]
```

Prep data
```{r}
# Remove any NA values (LFC==NA and geneName==NA)
protGSEA <- protLFQ_parse[LFQ_FetB.vs.WT != "NA"][geneName != "NA"]
# Add column of entrez IDs
length(intersect(protGSEA$geneName, keys(org.Hs.eg.db, keytype="SYMBOL")))
length(intersect(protGSEA$geneName, keys(org.Hs.eg.db, keytype="ALIAS")))
entrezMap <- AnnotationDbi::select(org.Hs.eg.db, keys=protGSEA$geneName, keytype = "SYMBOL", columns="ENTREZID") #if row has multiple gene names, only use first to map
idx <- match(protGSEA$geneName, entrezMap$SYMBOL)
#View(data.frame(protGSEA$geneName, entrezMap$SYMBOL[idx])) #sanity check
protGSEA$Entrez <- entrezMap$ENTREZID[idx]

```

###################################################### 
ENRICHMENT OF ALL INTERACTORS
######################################################
```{r}
set.seed(18)
interactome.GO_MF <- enrichGO(gene = na.omit(protGSEA$geneName),
         #universe = na.omit(cluster.gene.map$geneName),
         keyType = 'SYMBOL',
         OrgDb = org.Hs.eg.db,
         minGSSize = 10,
         ont= "MF",
         pAdjustMethod = "BH",
         pvalueCutoff = 0.01,
         qvalueCutoff = 0.01) %>% clusterProfiler::simplify(cutoff = 0.6, by="p.adjust", select_fun=min)
View(data.table(interactome.GO_MF@result)[p.adjust <= 0.01 ,])

set.seed(18)
interactome.GO_BP <- enrichGO(gene = na.omit(protGSEA$geneName),
         #universe = na.omit(cluster.gene.map$geneName),
         keyType = 'SYMBOL',
         OrgDb = org.Hs.eg.db,
         minGSSize = 10,
         ont= "BP",
         pAdjustMethod = "BH",
         pvalueCutoff = 0.01,
         qvalueCutoff = 0.01) %>% clusterProfiler::simplify(cutoff = 0.6, by="p.adjust", select_fun=min)
View(data.table(interactome.GO_BP@result)[p.adjust <= 0.01 ,])

set.seed(18)
interactome.GO_CC <- enrichGO(gene = na.omit(protGSEA$geneName),
         #universe = na.omit(cluster.gene.map$geneName),
         keyType = 'SYMBOL',
         OrgDb = org.Hs.eg.db,
         minGSSize = 10,
         ont= "CC",
         pAdjustMethod = "BH",
         pvalueCutoff = 0.01,
         qvalueCutoff = 0.01) %>% clusterProfiler::simplify(cutoff = 0.6, by="p.adjust", select_fun=min)
View(data.table(interactome.GO_CC@result)[p.adjust <= 0.01 ,])

set.seed(18)
interactome.KEGG <- enrichKEGG(gene = na.omit(protGSEA$Entrez),
                            #universe = na.omit(cluster.gene.map$Entrez),
                            organism = "hsa",
                            minGSSize = 10,
                            pvalueCutoff = 0.05,
                            pAdjustMethod = "BH")
View(data.table(interactome.KEGG@result)[p.adjust <= 0.01 ,])

set.seed(18)
interactome.Reactome <- enrichPathway(gene = na.omit(protGSEA$Entrez),
                                   #universe = na.omit(cluster.gene.map$Entrez),
                                   organism = "human",
                                   minGSSize = 10,
                                   pvalueCutoff = 0.05,
                                   pAdjustMethod = 'BH')
View(data.table(interactome.Reactome@result)[p.adjust <= 0.01 ,])
```



###################################################### 
DATA VISUALIZATION
######################################################
```{r prep data}
interaction.data <- protLFQ[, c("TRAb_R1", "TRAb_R2", "TRAb_R3", "WTAb_R1", "WTAb_R2", "WTAb_R3")] %>%
  setNames(c("NAFLD_R1", "NAFLD_R2", "NAFLD_R3", "WT_R1", "WT_R2", "WT_R3")) %>%
  as.matrix() %>%
  BBmisc::setRowNames(seq(1:nrow(protLFQ)))
cellLine.labels <- as.factor(c(rep("NAFLD",3), rep("WT", 3)))
cellLine_col <- brewer.pal(4,"Dark2")[as.numeric(cellLine.labels)]
```

PCA
```{r PCA analysis}
# PCA
res.pca <- prcomp(t(interaction.data), scale=F)

# visualize eigenvalues/variances
fviz_screeplot(res.pca, addlabels=T, ylim = c(0, 100))

# plot individuals on first two components
fviz_pca_ind(res.pca,
             geom.ind = c("text", "point"),
             pointsize = 3,
             mean.point = FALSE,
             show.legend=FALSE,
             habillage = as.factor(cellLine.labels),
             palette = "Dark2",
             legend.title = "Cell Line",
             repel = TRUE     # Avoid text overlapping
) + theme(text=element_text(size=4), axis.text = element_text(size = 10), axis.title = element_text(size = 10), legend.text = element_text(size = 10), legend.title = element_text(size = 10), plot.title = element_text(size = 20))

## Variable contributions to the principal axes
# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 5)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 5)

# Graph of variables
fviz_pca_var(res.pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
)

```

check outlierness of samples
```{r}
protLFQ.samples <- readxl::read_excel("Input/Summary_Results_HEP.xlsx", sheet="Perseus") %>% data.table() %>% select(c("TRAb_R1", "TRAb_R2", "TRAb_R3", "WTAb_R1", "WTAb_R2", "WTAb_R3"))
colSums(protLFQ.samples)
plot(x=1:6, y=colSums(protLFQ.samples))
```
Pearson correlation among biological replicates
```{r}
cor.test(protLFQ.samples$TRAb_R1, protLFQ.samples$TRAb_R2, method=c("pearson"))[c("estimate", "p.value")] %>% as.data.frame() %>% `rownames<-`("T1.T2") %>%
  rbind(cor.test(protLFQ.samples$TRAb_R1, protLFQ.samples$TRAb_R3, method=c("pearson"))[c("estimate", "p.value")] %>% as.data.frame() %>% `rownames<-`("T1.T3")) %>%
  rbind(cor.test(protLFQ.samples$TRAb_R2, protLFQ.samples$TRAb_R3, method=c("pearson"))[c("estimate", "p.value")] %>% as.data.frame() %>% `rownames<-`("T2.T3"))

cor.test(protLFQ.samples$WTAb_R1, protLFQ.samples$WTAb_R2, method=c("pearson"))[c("estimate", "p.value")] %>% as.data.frame() %>% `rownames<-`("W1.W2") %>%
  rbind(cor.test(protLFQ.samples$WTAb_R1, protLFQ.samples$WTAb_R3, method=c("pearson"))[c("estimate", "p.value")] %>% as.data.frame() %>% `rownames<-`("W1.W3")) %>%
  rbind(cor.test(protLFQ.samples$WTAb_R2, protLFQ.samples$WTAb_R3, method=c("pearson"))[c("estimate", "p.value")] %>% as.data.frame() %>% `rownames<-`("W2.W3"))
```

Hierarchical clustering heatmap
```{r}
hclustfunc <-function(x) hclust(x, method="complete")
distfunc <- function(x) dist(x, method="euclidean")
scale.data <- scale(t(interaction.data))
d <- distfunc(t(scale.data))
fit <- hclustfunc(d)
groups <- cutree(fit, k=2)
cols <- brewer.pal(4, "Dark2")[3:4]

png("Figures/hmap_interactome.png", width=850, height=1000)
par(cex.main=3)
hmap <- gplots::heatmap.2(t(scale.data),
                  scale = "none",
                  hclust=hclustfunc,
                  distfun=distfunc,
                  main="Interactomics",
                  trace="none", 
                  key=T,
                  labRow = "",
                  dendrogram = c("column"),
                  denscol = "black",
                  ColSideColors = cellLine_col,
                  RowSideColors = cols[groups],
                  key.xlab = "Interaction Z-Score",
                  key.ylab = "",
                  key.title = "Color Key",
                  key.size = 2,
                  key.par = list(cex.lab=2.5, cex.main=2.5, cex.axis=2.5, cex.names=2.5, mar=c(6,5.5,6,0)),
                  key.xtickfun = function(){
                    mtext("Count",cex=2, adj=0.5, side=2, line=3, padj=0)
                  },
                  srtCol = 45,
                  cexCol=3,
                  offsetCol=0.05,
                  mar=c(11,5))
legend(y=0.7, x=-0.05, xpd=TRUE, legend = c("Cluster 1", "Cluster 2"), fill = unique(cols[groups]), cex=2.2)
legend(y=0.55, x=-0.05, xpd=TRUE, legend = c("NAFLD", "WT"), fill = unique(cellLine_col), cex=2.2)
dev.off()

#par(mar = c(8, 8, 8, 8)) # Set the margin on all sides to 6
#eval(hmap$call)
#legend(y=1.6, x=1, xpd=TRUE, legend = c("Cluster 1", "Cluster 2"), fill = unique(cols[groups]), cex=1)
#legend(y=1.3, x=1, xpd=TRUE, legend = c("NAFLD", "WT"), fill = unique(cellLine_col), cex=1)
```

OLD gene cluster trend -- very noisy
```{r}
gene_cluster <- tibble::enframe(groups) %>% setNames(c("gene", "cluster")) %>%
  inner_join(data.frame(gene=colnames(scale.data), t(scale.data))) %>% setcolorder(c("gene", "cluster", rownames(scale.data)[hmap$colInd]))
#convert wide to long
gene_cluster.long <- gather(gene_cluster, sample, LFQ, NAFLD_R3:WT_R1, factor_key=T)
ggplot(gene_cluster.long, aes(sample, LFQ)) +
  geom_line(aes(group=gene), alpha = 0.3) +
  geom_line(stat="summary", fun="median", colour="brown", size=1.5, aes(group=1)) +
  facet_grid(cols=vars(cluster))
```

Generate trend summaries -- plot using summary data to eliminate noise
```{r}
library(dplyr)
library(Rmisc)
library(ggplot2)

gene_cluster.summary <- ddply(gene_cluster.long, c("cluster", "sample"), summarise,
                              mean=CI(LFQ, ci=0.95)['mean'],
                              sd=sd(LFQ),
                              se=plotrix::std.error(LFQ),
                              uci=CI(LFQ, ci=0.95)['upper'],
                              lci=CI(LFQ, ci=0.95)['lower'])
cluster.1 <- gene_cluster.summary[gene_cluster.summary$cluster==1 ,] %>% mutate(index=1:6)
cluster.2 <- gene_cluster.summary[gene_cluster.summary$cluster==2 ,] %>% mutate(index=1:6)
```

NEW gene cluster trend -- using shaded ribbon to show spread
```{r}
trend.ribbon <- function(summary.data) {
  ggplot(summary.data, aes(index, mean)) +
  geom_line(color="firebrick", size=1) +
  geom_ribbon(aes(ymin=mean-se, ymax=mean+se), alpha=0.1, color="black", linetype="dotted")
}

trend.ribbon(cluster.1)
trend.ribbon(cluster.2)
```

NEW gene cluster trend -- using error bars wth 95 CI
```{r}
trend.errbar <- function(summary.data) {
  ggplot(summary.data, aes(sample, mean, group=cluster)) +
  geom_point() +
  geom_line(color="firebrick", size=2) +
  geom_errorbar(aes(ymin=lci, ymax=uci), width=0.1) +
  scale_x_discrete(labels= as.character(summary.data$sample)) +
    theme(text=element_text(size=20), axis.text.x=element_text(angle = -45, hjust = 0))
}

trend.errbar(cluster.1)
trend.errbar(cluster.2)
```


Cluster 1
```{r}
#length(setdiff(names(groups), keys(org.Hs.eg.db, keytype="ALIAS")))
## replace protein group numeric id with gene names
cluster.gene.map <- full_join(protGSEA, gene_cluster[, c("gene", "cluster")], by=c("id"="gene"))

set.seed(18)
GO.MF_cluster1 <- enrichGO(gene = na.omit(cluster.gene.map$geneName[cluster.gene.map$cluster==1]),
         universe = na.omit(cluster.gene.map$geneName),
         keyType = 'SYMBOL',
         OrgDb = org.Hs.eg.db,
         minGSSize = 8,
         ont= "MF",
         pAdjustMethod = "BH",
         pvalueCutoff = 0.05,
         qvalueCutoff = 0.1) %>% clusterProfiler::simplify(cutoff = 0.8, by="p.adjust", select_fun=min) #%>% clusterProfiler::simplify(cutoff = 0.8, by="p.adjust", select_fun=min)
View(data.table(GO.MF_cluster1@result)[p.adjust <= 0.1 ,])

set.seed(18)
GO.BP_cluster1 <- enrichGO(gene = na.omit(cluster.gene.map$geneName[cluster.gene.map$cluster==1]),
         universe = na.omit(cluster.gene.map$geneName),
         keyType = 'SYMBOL',
         OrgDb = org.Hs.eg.db,
         minGSSize = 8,
         ont= "BP",
         pAdjustMethod = "BH",
         pvalueCutoff = 0.05,
         qvalueCutoff = 0.1) %>% clusterProfiler::simplify(cutoff = 0.8, by="p.adjust", select_fun=min)
View(data.table(GO.BP_cluster1@result)[p.adjust <= 0.1 ,])

set.seed(18)
GO.CC_cluster1 <- enrichGO(gene = na.omit(cluster.gene.map$geneName[cluster.gene.map$cluster==1]),
         universe = na.omit(cluster.gene.map$geneName),
         keyType = 'SYMBOL',
         OrgDb = org.Hs.eg.db,
         minGSSize = 8,
         ont= "CC",
         pAdjustMethod = "BH",
         pvalueCutoff = 0.05,
         qvalueCutoff = 0.1) %>% clusterProfiler::simplify(cutoff = 0.8, by="p.adjust", select_fun=min)
View(data.table(GO.CC_cluster1@result)[p.adjust <= 0.1 ,])

######### KEGG and Reactome enrichment

set.seed(18)
KEGG_cluster1 <- enrichKEGG(gene = na.omit(cluster.gene.map$Entrez[cluster.gene.map$cluster==1]),
                            universe = na.omit(cluster.gene.map$Entrez),
                            organism = "hsa",
                            minGSSize = 8,
                            pvalueCutoff = 0.05,
                            pAdjustMethod = "BH")
View(data.table(KEGG_cluster1@result)[p.adjust <= 0.1 ,])

set.seed(18)
Reactome_cluster1 <- enrichPathway(gene = na.omit(cluster.gene.map$Entrez[cluster.gene.map$cluster==1]),
                                   universe = na.omit(cluster.gene.map$Entrez),
                                   organism = "human",
                                   minGSSize = 8,
                                   pvalueCutoff = 0.05,
                                   pAdjustMethod = 'BH')
View(data.table(Reactome_cluster1@result)[p.adjust <= 0.1 ,])
```

Cluster 2
```{r}
set.seed(18)
GO.MF_cluster2 <- enrichGO(gene = na.omit(cluster.gene.map$geneName[cluster.gene.map$cluster==2]),
         universe = na.omit(cluster.gene.map$geneName),
         keyType = 'SYMBOL',
         OrgDb = org.Hs.eg.db,
         minGSSize = 8,
         ont= "MF",
         pAdjustMethod = "BH",
         pvalueCutoff = 0.05,
         qvalueCutoff = 0.1) %>% clusterProfiler::simplify(cutoff = 0.8, by="p.adjust", select_fun=min)
View(data.table(GO.MF_cluster2@result)[p.adjust <= 0.1 ,])

set.seed(18)
GO.BP_cluster2 <- enrichGO(gene = na.omit(cluster.gene.map$geneName[cluster.gene.map$cluster==2]),
         universe = na.omit(cluster.gene.map$geneName),
         keyType = 'SYMBOL',
         OrgDb = org.Hs.eg.db,
         minGSSize = 8,
         ont= "BP",
         pAdjustMethod = "BH",
         pvalueCutoff = 0.05,
         qvalueCutoff = 0.1) %>% clusterProfiler::simplify(cutoff = 0.8, by="p.adjust", select_fun=min)
View(data.table(GO.BP_cluster2@result)[p.adjust <= 0.1 ,])

set.seed(18)
GO.CC_cluster2 <- enrichGO(gene = na.omit(cluster.gene.map$geneName[cluster.gene.map$cluster==2]),
         universe = na.omit(cluster.gene.map$geneName),
         keyType = 'SYMBOL',
         OrgDb = org.Hs.eg.db,
         minGSSize = 8,
         ont= "CC",
         pAdjustMethod = "BH",
         pvalueCutoff = 0.05,
         qvalueCutoff = 0.1) %>% clusterProfiler::simplify(cutoff = 0.8, by="p.adjust", select_fun=min)
View(data.table(GO.CC_cluster2@result)[p.adjust <= 0.1 ,])

######### KEGG and Reactome enrichment

set.seed(18)
KEGG_cluster2 <- enrichKEGG(gene = na.omit(cluster.gene.map$Entrez[cluster.gene.map$cluster==2]),
                            universe = na.omit(cluster.gene.map$Entrez),
                            organism = "hsa",
                            minGSSize = 8,
                            pvalueCutoff = 0.05,
                            pAdjustMethod = "BH")
View(data.table(KEGG_cluster2@result)[p.adjust <= 0.1 ,])

set.seed(18)
Reactome_cluster2 <- enrichPathway(gene = na.omit(cluster.gene.map$Entrez[cluster.gene.map$cluster==2]),
                                   universe = na.omit(cluster.gene.map$Entrez),
                                   organism = "human",
                                   minGSSize = 8,
                                   pvalueCutoff = 0.05,
                                   pAdjustMethod = 'BH')
View(data.table(Reactome_cluster2@result)[p.adjust <= 0.1 ,])
```


```{r}

gplots::heatmap.2(interaction.data,
                  scale = "none",
                  main="Interactomics Heatmap",
                  trace="none", key=T,
                  labRow = "",
                  denscol = "grey",
                  ColSideColors = cellLine_col,
                  #col=colorRampPalette(c("darkblue","white","darkred"))(100),
                  col=rainbow(100),
                  srtCol = 45,
                  cexCol=1.3,
                  offsetCol=0.05)
legend(y=0.8, x=-0.1, xpd=TRUE,
    legend = c("NAFLD", "WT"), # category labels
    fill = unique(cellLine_col),
    cex=0.8
)
dev.off()
```

```{r}
interaction.data <- data.table(protLFQ[, c("TRAb_R1", "TRAb_R2", "TRAb_R3", "WTAb_R1", "WTAb_R2", "WTAb_R3")])
interaction.data[, FetB.mean := rowMeans(interaction.data[, c("TRAb_R1", "TRAb_R2", "TRAb_R3")])][, WT.mean := rowMeans(interaction.data[, c("WTAb_R1", "WTAb_R2", "WTAb_R3")])][, LFC := FetB.mean - WT.mean][,old_LFC := protLFQ$`Welch's T-test Difference LFQ intensity TRAb_LFQ intensity WTAb`]
bob <- data.table(matrixTests::row_t_welch(x=interaction.data[, c("TRAb_R1", "TRAb_R2", "TRAb_R3")], y=interaction.data[, c("WTAb_R1", "WTAb_R2", "WTAb_R3")], alternative = "two.sided"))
bob[, FDR := p.adjust(pvalue, method="BH")]
```

Volcano plot
```{r}
lab_italics <- paste0("italic('", protLFQ_trim$geneName, "')")
selectLab_italics <- paste0(
    "italic('",
    protLFQ_trim$geneName[protLFQ_trim$qvalue <= 0.1],
    "')")
png("Figures/volcano_interactome.png", width=700, height=900)
EnhancedVolcano::EnhancedVolcano(protLFQ_trim, 
                                 lab=lab_italics, 
                                 parseLabels=T,
                                 boxedLabels = T,
                                 x= "LFQ_FetB.vs.WT", 
                                 y="qvalue", 
                                 axisLabSize = 30,
                                 pCutoff=0.1, 
                                 FCcutoff = 1.5,
                                 labSize = 10,
                                 labCol="black",
                                 labFace="bold",
                                 title = "Interactomics Volano Plot",
                                 titleLabSize = 30,
                                 subtitle = 'NAFLD vs. WT',
                                 subtitleLabSize = 28,
                                 ylim=c(0,3),
                                 pointSize = c(ifelse(protLFQ_trim$qvalue<=0.1, 10, 5)),
                                 caption = bquote(~Log[2]~ "fold change cutoff: 1.5; FDR cutoff: 0.1"),
                                 captionLabSize = 25,
                                 legendPosition = "bottom",
                                 legendLabels = c("NS", ~Log[2]~"FC", "FDR", ~Log[2]~"FC and FDR"),
                                 legendLabSize = 28,
                                 colAlpha = 0.9,
                                 #colGradient = c('red3', 'royalblue'),
                                 drawConnectors = T,
                                 widthConnectors = 0,
                                 ylab = expression("-Log"[10]*" FDR")
                                 )
dev.off()
```


###################################################### 
GSEA
######################################################

```{r}
# Extract and name LFC
LFC_entrez <- protGSEA$LFQ_FetB.vs.WT[!is.na(protGSEA$Entrez)] %>% setNames(protGSEA$Entrez[!is.na(protGSEA$Entrez)]) %>% sort(decreasing = T)
LFC_geneName <- protGSEA$LFQ_FetB.vs.WT[!is.na(protGSEA$geneName)] %>% setNames(protGSEA$geneName[!is.na(protGSEA$geneName)]) %>% sort(decreasing=T)

# Remove any duplicates
LFC_entrez <- LFC_entrez[which(duplicated(names(LFC_entrez))==F)]
LFC_geneName <- LFC_geneName[which(duplicated(names(LFC_geneName))==F)]
```


GSEA using gene sets from KEGG
```{r}
set.seed(18)
gseaKEGG <- gseKEGG(geneList = LFC_entrez,
        organism = "hsa",
        minGSSize = 10,
        pvalueCutoff = 0.01,
        pAdjustMethod = "BH",
        verbose = F) %>% setReadable(OrgDb=org.Hs.eg.db, "ENTREZID")
View(data.table(gseaKEGG@result)[p.adjust <= 0.05 ,])
```

GSEA using gene sets from GO Cellular Component
```{r}
set.seed(18)
gseaGO_CC <- gseGO(geneList = LFC_geneName,
                   OrgDb = org.Hs.eg.db,
                   keyType="SYMBOL",
                   ont = 'CC',
                   minGSSize = 10,
                   pvalueCutoff = 0.01,
                   pAdjustMethod = 'BH',
                   verbose = F) %>% clusterProfiler::simplify(cutoff = 0.8, by="p.adjust", select_fun=min)
View(data.table(gseaGO_CC@result)[p.adjust <= 0.05 ,])
```

GSEA using gene sets from GO Biological Process
```{r}
set.seed(18)
gseaGO_BP <- gseGO(geneList = LFC_geneName,
                   OrgDb = org.Hs.eg.db,
                   keyType="SYMBOL",
                   ont = 'BP',
                   minGSSize = 10,
                   pvalueCutoff = 0.05,
                   pAdjustMethod = 'BH',
                   verbose = F) %>% clusterProfiler::simplify(cutoff = 0.8, by="p.adjust", select_fun=min)
View(data.table(gseaGO_BP@result)[p.adjust <= 0.05 ,])

```

GSEA using gene sets from GO Molecular Function
```{r}
set.seed(18)
gseaGO_MF <- gseGO(geneList = LFC_geneName,
                   OrgDb = org.Hs.eg.db,
                   keyType="SYMBOL",
                   ont = 'MF',
                   minGSSize = 10,
                   pvalueCutoff = 0.05,
                   pAdjustMethod = 'BH',
                   verbose = F) %>% clusterProfiler::simplify(cutoff = 0.8, by="p.adjust", select_fun=min)
View(data.table(gseaGO_MF@result)[p.adjust <= 0.05 ,])

```

GSEA using gene sets from Reactome
```{r}
set.seed(18)
gseaReactome <- gsePathway(geneList = LFC_entrez,
                           organism = "human",
                           minGSSize = 10,
                           pvalueCutoff = 0.05,
                           pAdjustMethod = 'BH',
                           verbose = F) %>% setReadable(OrgDb=org.Hs.eg.db, "ENTREZID")
View(data.table(gseaReactome@result)[p.adjust <= 0.05 ,])
```

```{r}
c2 <- read.gmt("Input/c2.all.v7.2.entrez.gmt")
set.seed(18)
msig_c2 <- GSEA(LFC_entrez[names(LFC_entrez) %in% c2$gene], TERM2GENE=c2, verbose=FALSE, pvalueCutoff = 0.05, pAdjustMethod = 'BH') %>% setReadable(OrgDb=org.Hs.eg.db, "ENTREZID")
View(data.table(msig_c2@result)[p.adjust <= 0.05 ,])
```

```{r}
c5 <- read.gmt("Input/c5.all.v7.2.entrez.gmt")
set.seed(18)
msig_c5 <- GSEA(LFC_entrez[names(LFC_entrez) %in% c5$gene], TERM2GENE=c5, verbose=FALSE, pvalueCutoff = 0.05, pAdjustMethod = 'BH') %>% setReadable(OrgDb=org.Hs.eg.db, "ENTREZID")
View(data.table(msig_c5@result)[p.adjust <= 0.05 ,])
```

save GSEA results in excel workbook to cherry pick from
```{r}
allGSEA <- list("KEGG"=data.table(gseaKEGG@result)[p.adjust <= 0.05 ,], "Reactome"=data.table(gseaReactome@result)[p.adjust <= 0.05 ,], "GO_CC"=data.table(gseaGO_CC@result)[p.adjust <= 0.05 ,], "GO_BP"=data.table(gseaGO_BP@result)[p.adjust <= 0.05 ,], "GO_MF"=data.table(gseaGO_MF@result)[p.adjust <= 0.05 ,], "MSigDB_c2"=data.table(msig_c2@result)[p.adjust <= 0.05 ,], "MSigDB_c5"=data.table(msig_c5@result)[p.adjust <= 0.05 ,])
write.xlsx(allGSEA, "Output/gseaResults_interactomeCaressa.xlsx")
saveRDS(allGSEA, "Output/allGSEA_interactomeCaressa.RDS")
```

Plot interesting results
```{r}
# Load cherry picked interesting results
gsea_all <- readxl::read_excel("Output/gseaResults_interactomeCaressa_cherryPick.xlsx", sheet="cherryPick")
numCoreEnrich <- lapply(gsea_all$core_enrichment, function(x) {strsplit(x,'/')})
num <- lapply(numCoreEnrich, function(x) {length(unlist(x))})
gsea_all$coreEnrichment <- unlist(num)
gsea_all$Description <- tolower(gsea_all$Description)

subset.interesting <- gsea_all
subset.interesting$Description <- factor(subset.interesting$Description, levels = subset.interesting$Description[order(subset.interesting$NES)])

# color terms related to secretory pathway
term.color <- ifelse(subset.interesting$Tag == "secPathway", "red", "gray25")

## dotplot
ggplot(subset.interesting, # you can replace the numbers to the row number of pathway of your interest
       aes(x = NES, y = Description)) +
  geom_point(aes(size = coreEnrichment, color = p.adjust)) +
  theme_bw(base_size = 19) +
  scale_colour_gradient(limits=c(0, 0.05), low="red", high="blue") +
  ylab(NULL) + 
  theme(axis.text.y = element_text(colour=term.color), legend.title=element_text(size=14), axis.title = element_text(size = 14)) +
  ggtitle("GSEA of Fetuin B Interactome")

## barplot
ggplot(subset.interesting, # you can replace the numbers to the row number of pathway of your interest
       aes(x = NES, y = Description)) +
  geom_col(aes(fill = p.adjust)) +
  theme_bw(base_size = 20) +
  scale_fill_gradient(limits=c(0, 0.05), low="red", high="blue") +
  ylab(NULL) + 
  theme(axis.text.y = element_text(colour=term.color)) +
  ggtitle("GSEA of Fetuin B Interactome")
```

Lolipop
```{r}
ross <- gsea_all %>% arrange(NES) 
ross %>%
  mutate(Description=factor(Description, levels=Description)) %>%
  ggplot( aes(x=Description, y=NES, color=database)) +
    geom_segment( aes(xend=Description, yend=0), color = "grey50") +
    geom_point(size=50*ross$p.adjust) +
    coord_flip() +
    theme_bw() +
    xlab("")
rm(ross)

roman <- gsea_all[!(gsea_all$database %in% c("MSigDB_c2", "MSigDB_c5")) ,] %>% arrange(NES) 
roman %>%
  mutate(Description=factor(Description, levels=Description)) %>%
  ggplot( aes(x=Description, y=NES, color=database)) +
    geom_segment( aes(xend=Description, yend=0), color = "grey50") +
    geom_point(size=50*roman$p.adjust) +
    coord_flip() +
    theme_bw() +
    xlab("")
rm(roman)
```



Hypergeometric enrichment of sig diff interactions
```{r}
sigInteractions <- protGSEA[qvalue <= 0.25 & abs(LFQ_FetB.vs.WT) >= log2(1.5) ,][order(LFQ_FetB.vs.WT)]

# GO classification
ggo <- groupGO(gene=sigInteractions$geneName[sigInteractions$LFQ_FetB.vs.WT >0],
               keyType = "SYMBOL",
               OrgDb = org.Hs.eg.db,
               ont ="BP",
               level=3)
# GO hypergeometric enrichment
set.seed(18)
ego <- enrichGO(gene=sigInteractions$geneName[sigInteractions$LFQ_FetB.vs.WT > 0],
                keyType = "SYMBOL",
                universe=protGSEA$geneName,
                OrgDb = org.Hs.eg.db,
                ont="CC",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.05) %>% simplify(cutoff = 0.8, by="p.adjust", select_fun=min)

# KEGG hypergeometric enrichment
set.seed(18)
kk <- enrichKEGG(gene=sigInteractions$Entrez[sigInteractions$LFQ_FetB.vs.WT > 0],
                 universe=protGSEA$Entrez,
                 organism='hsa',
                 pvalueCutoff = 0.05)

# KEGG module hypergeometric enrichment
set.seed(18)
kkm <- enrichMKEGG(gene=sigInteractions$Entrez[sigInteractions$LFQ_FetB.vs.WT > 0],
                  universe=protGSEA$Entrez,
                 organism='hsa',
                 pvalueCutoff = 0.05)

# Reactome
set.seed(18)
epath <- enrichPathway(gene=sigInteractions$Entrez[sigInteractions$LFQ_FetB.vs.WT > 0],
                       universe=protGSEA$Entrez,
                pvalueCutoff = 0.05)
```









###################################################### 
GSEA USIG ORGANELLE ASSIGNMENTS: KRAHMER ET AL 2018 
######################################################

Load organelle assignment data
```{r}
organelle.Krahmer <- readxl::read_excel("Input/Krahmer_2018_S3.xlsx", sheet="Organelleassignaments")
```

```{r}
organelle.HFD3 <- data.table(organelle.Krahmer)[, c("Gene names","Main organelle_HFD3", "Second organelle_HFD3", "Alpha_HFD3", "Corr_HFD3")] %>% setNames(c("geneName", "HFD3_main", "HFD3_second", "HFD3_alpha", "HFD3_corr"))
#note: can't include protein names because number of proteins names does NOT equal number of gene names for split
organelle.HFD3_parse <- mutate(organelle.HFD3, id=1:nrow(organelle.HFD3)) %>% tidyr::separate_rows(geneName, sep=";") %>% data.table()
term2gene_HFD3.main <- organelle.HFD3_parse[, c("HFD3_main", "geneName")][, geneName := toupper(geneName)] %>% na.omit()
term2gene_HFD3.second <- organelle.HFD3_parse[, c("HFD3_second", "geneName")][, geneName := toupper(geneName)] %>% na.omit()
```

```{r}
organelle.HFD12 <- data.table(organelle.Krahmer)[, c("Gene names","Main organelle_HFD12", "Second organelle_HFD12", "Alpha_HFD12", "Corr_HFD12")] %>% setNames(c("geneName", "HFD12_main", "HFD12_second", "HFD12_alpha", "HFD12_corr"))
#note: can't include protein names because number of proteins names does NOT equal number of gene names for split
organelle.HFD12_parse <- mutate(organelle.HFD12, id=1:nrow(organelle.HFD12)) %>% tidyr::separate_rows(geneName, sep=";") %>% data.table()
term2gene_HFD12.main <- organelle.HFD12_parse[, c("HFD12_main", "geneName")][, geneName := toupper(geneName)] %>% na.omit()
term2gene_HFD12.second <- organelle.HFD12_parse[, c("HFD12_second", "geneName")][, geneName := toupper(geneName)] %>% na.omit()
```

```{r}
length(setdiff(names(LFC_geneName), term2gene_HFD3.main$geneName))
set.seed(18)
gseaKrahmer.HFD3main <- GSEA(LFC_geneName[names(LFC_geneName) %in% term2gene_HFD3.main$geneName], TERM2GENE=term2gene_HFD3.main, verbose=FALSE, pvalueCutoff = 0.05, pAdjustMethod = 'BH')
set.seed(18)
gseaKrahmer.HFD3second <- GSEA(LFC_geneName[names(LFC_geneName) %in% term2gene_HFD3.second$geneName], TERM2GENE=term2gene_HFD3.second, verbose=FALSE, pvalueCutoff = 0.05, pAdjustMethod = 'BH')
```

```{r}
length(setdiff(names(LFC_geneName), toupper(organelle.HFD12_parse$geneName)))
set.seed(18)
gseaKrahmer.HFD12main <- GSEA(LFC_geneName[names(LFC_geneName) %in% term2gene_HFD12.main$geneName], TERM2GENE=term2gene_HFD12.main, verbose=FALSE, pvalueCutoff = 0.05, pAdjustMethod = 'BH')
set.seed(18)
gseaKrahmer.HFD12second <- GSEA(LFC_geneName[names(LFC_geneName) %in% term2gene_HFD12.second$geneName], TERM2GENE=term2gene_HFD12.second, verbose=FALSE, pvalueCutoff = 0.05, pAdjustMethod = 'BH')
```

```{r}
golgi.depletion <- strsplit(gseaKrahmer.HFD3main@result$core_enrichment,split='/', fixed=TRUE) %>% unlist()
View(protGSEA[protGSEA$geneName %in% golgi.depletion ,])
```

###################################################### 
KRAHMER OUTLIER TEST -- RELOCALIZATION 
######################################################

```{r}
outlierTest3 <- readxl::read_excel("Input/Krahmer_2018_S5.xlsx", sheet="outliertest3") %>% data.table()
outlierTest12 <- readxl::read_excel("Input/Krahmer_2018_S5.xlsx", sheet="outliertest12") %>% data.table()
outlierTest3.sig <- outlierTest3[`overall Hit relocalization test`=="+", `Gene names`] %>% data.table() %>% setNames("geneName") %>% tidyr::separate_rows(geneName, sep=";") %>% data.table()
outlierTest12.sig <- outlierTest12[`Overall Hit relocalization test`=="+", `Gene names`] %>% data.table() %>% setNames("geneName") %>% tidyr::separate_rows(geneName, sep=";") %>% data.table()
```

```{r}
outlier3.si
```



###################################################### 
LOOKING INTO FORMS OF VESICLE TRANSPORT 
######################################################

Looking into GO for vesicle specific terms
```{r}
library(GO.db)

length(intersect(protGSEA$geneName, keys(org.Hs.eg.db, keytype="SYMBOL")))
length(intersect(protGSEA$Entrez, keys(org.Hs.eg.db, keytype="ENTREZID")))
GOterms <- data.table(AnnotationDbi::select(org.Hs.eg.db, keys=protGSEA$geneName, columns = c("GO"), keytype = "SYMBOL"))
#only include CC terms
GOterms_CC <- GOterms[ONTOLOGY=="CC"]
setnames(GOterms_CC, "GO", "GOID")
# remove any duplicates
sum(duplicated(GOterms_CC))
GOterms_CC <- distinct(GOterms_CC)

#get corresponding GO terms
ID2term <- AnnotationDbi::select(GO.db, keys=GOterms_CC$GO, columns="TERM", keytype = "GOID") %>% distinct()
map_gene2term <- full_join(GOterms_CC, ID2term)
# check and remove any duplicates
sum(duplicated(map_gene2term))
map_gene2term <- distinct(map_gene2term)

secretionTerms <- c("exocytic vesicle", "secretory granule", "extracellular exosome", "autophagosome", "endosome to plasma membrane transport vesicle", "autolysosome", "endosome", "lysosome", "cytoplasmic vesicle", "clathrin-coated vesicle", "transport vesicle", "phagocytic vesicle", "extracellular vesicle", "exocytic vesicle", "amphisome", "multivesicular body") 

secretionGenes <- map_gene2term[map_gene2term$TERM %in% secretionTerms, c("TERM", "SYMBOL")] %>% inner_join(protGSEA, by=c("SYMBOL"="geneName")) %>% distinct() %>% arrange(LFQ_FetB.vs.WT)

secGenes_pos <- secretionGenes[LFQ_FetB.vs.WT > 0] %>% distinct()
secGenes_neg <- secretionGenes[LFQ_FetB.vs.WT < 0] %>% distinct()

gsea_vesicle <- GSEA(LFC_geneName[names(LFC_geneName) %in% secretionGenes$SYMBOL], TERM2GENE=secretionGenes[,c("TERM", "SYMBOL")], verbose=FALSE, pvalueCutoff = 0.1, pAdjustMethod = 'BH')

```

```{r}

```

