---
title: "R Notebook"
output: html_notebook
---

Clear packages
```{r}
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE)
```

Load packages
```{r}
library(dplyr)
library(data.table)
library(factoextra)
library(ggplot2)
```

Load pheno and count data
```{r}
#pheno data
pheno <- openxlsx::read.xlsx("Input/Susan_Annotations.xlsx", sheet="clean") %>%
  mutate(condition_state = paste0(condition, "_", state))

#count data
counts <- fread("Input/count_clean.csv")
count.matrix <- as.matrix(counts[, -1]) %>% `rownames<-`(counts$entrezID) %>% na.omit()
```

Load Cellfie results
```{r}
res.cellfie <- fread("CellFie_output/local_minmaxmean/score.csv") %>% setNames(colnames(counts)[-1])
res.cellfie.binary <- fread("CellFie_output/local_minmaxmean/score_binary.csv") %>% setNames(colnames(counts)[-1])
taskInfo <- fread("CellFie_output/local_minmaxmean/taskInfo.csv") %>%
  mutate(Pathway = case_when(stringr::str_detect(taskInfos1, "sec_") ~ 'Secretory',
         !stringr::str_detect(taskInfos1, "sec_") ~ 'Metabolism')) %>%
  setNames(c("taskID", "Task", "System", "Subsystem", "Pathway"))
res.cellfie.filter <- res.cellfie * res.cellfie.binary 
```

############################################################################
Normalization
############################################################################
Samples CHO-Core-8 & CHO-Core-16 appear to skew from the ref line
To normalize, mult/divide by the slope of the best fit line

Function
```{r}
viz.samples <- function(condition, data) {
  sample.names <- pheno$sample.id[pheno$condition_state==condition]
  sample.pairs <- t(combn(sample.names,2))
  data.plot <- apply(sample.pairs, 1, function(i)
    {data.frame(x=select(data, i[1]), y=select(data,i[2]), pair=paste0(i[1],i[2])) %>% setNames(c("x", "y", "pair"))}) %>%
    dplyr::bind_rows()
  ggplot(data.plot, aes(x,y)) +
    geom_point(aes(color=pair, alpha=0.7)) +
    geom_abline(intercept = 0, slope = 1) + 
    theme(text=element_text(size=14)) +
    ggtitle(condition)
  } 

```

Plot pre-normalization for Supp Material
```{r}
png("Figures/rProtein_b4Norm.png", width=1100, height=700)
ggpubr::ggarrange(viz.samples("A0_EXP", data=res.cellfie), 
          viz.samples("A0_STA", data=res.cellfie),
          viz.samples("A1_EXP", data=res.cellfie),
          viz.samples("A1_STA", data=res.cellfie),
          labels= c("A", "B", "C", "D"),
          ncol=2, nrow=2)
dev.off()
```

Normalization
```{r}
m.8 <- coef(lm(res.cellfie$`CHO-Core-1`~res.cellfie$`CHO-Core-8`))
m.16 <- coef(lm(res.cellfie$`CHO-Core-16`~res.cellfie$`CHO-Core-17`))

res.cellfie.norm <- res.cellfie %>% mutate("CHO-Core-8"=`CHO-Core-8`*m.8[2]) %>%
  mutate("CHO-Core-16"=`CHO-Core-16`/m.16[2])

## sec only
sec.cellfie <- mutate(res.cellfie.norm, Task=taskInfo$Task, Pathway=taskInfo$Pathway) %>% 
  filter(Pathway=="Secretory") %>% select(-Pathway)

## met only
met.cellfie <- mutate(res.cellfie.norm, Task=taskInfo$Task, Pathway=taskInfo$Pathway) %>% 
  filter(Pathway=="Metabolism") %>% select(-Pathway)

rm(m.8)
rm(m.16)
```

Plot post-normalization for Supp Material
```{r}
png("Figures/rProtein_afterNorm.png", width=1100, height=700)
ggpubr::ggarrange(viz.samples("A0_EXP", data=res.cellfie.norm),
                  viz.samples("A0_STA", data=res.cellfie.norm),
                  viz.samples("A1_EXP", data=res.cellfie.norm),
                  viz.samples("A1_STA", data=res.cellfie.norm),
                  labels= c("A", "B", "C", "D"),
                  ncol=2, nrow=2)
dev.off()
```

############################################################################
PCA
############################################################################

Function
```{r}
PCA.plots <- function(matrix, pheno){
  # log transform
  matrix <- log2(matrix+0.001)
  # remove constant/zero columns
  matrix <- matrix[matrixStats::rowSds(matrix) != 0 ,]
  # PCA
  res.pca <- prcomp(t(matrix), center=T, scale=T) #, scale=T, center=T
  # visualize eigenvalues/variances
  ross <- fviz_screeplot(res.pca, addlabels=T, ylim = c(0, 100))
  # plot individuals on first two components
  ian <- fviz_pca_ind(res.pca,
               geom.ind = c("text", "point"), #c("text", "point")
               pointsize = 3,
               mean.point = FALSE,
               show.legend=FALSE,
               habillage = as.factor(pheno$condition_state),
               #palette = "Dark2",
               legend.title = "condition",
               repel = TRUE     # Avoid text overlapping
               ) + theme(text=element_text(size=4), axis.text = element_text(size = 10), axis.title =
                           element_text(size = 10), legend.text = element_text(size = 10), legend.title =
                           element_text(size = 10), plot.title = element_text(size = 20))
  return(list(results=res.pca, screeplot=ross, pca=ian))
}
```

PCA using sec+met CellFie
```{r}
cellfie.matrix <- mutate(res.cellfie.norm, Task=taskInfo$Task) %>%
  filter(`CHO-Core-1` >0) %>% # filter out -1 tasks (unable to calculate due to missing gene info)
  as.matrix(rownames="Task")
PCA.cellfie <- PCA.plots(cellfie.matrix, pheno=pheno)

png("Figures/cellfie_PCAbiplot.png", width=620, height=480)
fviz_pca_biplot(PCA.cellfie$results,
                select.var = list(contrib = 4),
                label="var",
                labelsize = 6,
                #col.var = "black",
                geom.ind = c("text", "point"),
                pointsize = 4,
                mean.point = FALSE,
                show.legend=FALSE,
                habillage = as.factor(pheno$condition_state),
                legend.title = "condition",
                repel = TRUE) + theme(text=element_text(size=10), axis.text = element_text(size = 18), axis.title =
                           element_text(size = 18), legend.text = element_text(size = 18), legend.title =
                           element_text(size = 18), plot.title = element_text(size = 22)) +
  scale_shape_manual(values=c(19,15,17,18)) +
  # ADD ggforce's ellipses
  ggforce::geom_mark_ellipse(aes(fill = pheno$condition_state,
                        color = pheno$condition_state), alpha=0.2) + ylim (-12, 9) + 
  labs(title="Principal Component Analysis")
dev.off()

png("Figures/cellfie_PCA.png", width=620, height=480)
fviz_pca_ind(PCA.cellfie$results,
               geom.ind = c("point"), #c("text", "point")
               pointsize = 5,
               mean.point = F,
               show.legend=FALSE,
               habillage = as.factor(pheno$condition_state),
               legend.title = "condition",
               repel = TRUE     # Avoid text overlapping
               ) + theme(text=element_text(size=10), axis.text = element_text(size = 18), axis.title =
                           element_text(size = 18), legend.text = element_text(size = 18), legend.title =
                           element_text(size = 18), plot.title = element_text(size = 22)) +
  scale_shape_manual(values=c(19,15,17,18)) +
  # ADD ggforce's ellipses
  ggforce::geom_mark_ellipse(aes(fill = pheno$condition_state,
                        color = pheno$condition_state)) + ylim (-12, 9) + 
  labs(title="Principal Component Analysis")
dev.off()
```

PCA using sec CellFie only
```{r}
### sec Cellfie
cellfie.matrix.sec <- filter(sec.cellfie,`CHO-Core-1` >0) %>%
  as.matrix(rownames="Task")
PCA.cellfie.sec <- PCA.plots(cellfie.matrix.sec, pheno=pheno)

png("Figures/secCellfie_biplot.png", width=620, height=480)
fviz_pca_biplot(PCA.cellfie.sec$results,
                select.var = list(contrib = 4),
                labelsize = 5,
                col.var="gray59",
               geom.ind = c("point"), #c("text", "point")
               pointsize = 5,
               mean.point = F,
               show.legend=FALSE,
               habillage = as.factor(pheno$condition_state),
               #palette = "Dark2",
               legend.title = "condition",
               repel = TRUE     # Avoid text overlapping
               ) + theme(text=element_text(size=10), axis.text = element_text(size = 18), axis.title =
                           element_text(size = 18), legend.text = element_text(size = 18), legend.title =
                           element_text(size = 18), plot.title = element_text(size = 22)) +
  labs(title="Principal Component Analysis", subtitle = "Secretory Pathway") +
  scale_shape_manual(values=c(19,15,17,18)) +
  theme(plot.subtitle = element_text(size=15, face = "italic"))
dev.off()

png("Figures/secCellfie_PCA.png", width=620, height=480)
fviz_pca_ind(PCA.cellfie.sec$results,
               geom.ind = c("point"), #c("text", "point")
               pointsize = 5,
               mean.point = F,
               show.legend=FALSE,
               habillage = as.factor(pheno$condition_state),
               legend.title = "condition",
               repel = TRUE     # Avoid text overlapping
               ) + theme(text=element_text(size=10), axis.text = element_text(size = 18), axis.title =
                           element_text(size = 18), legend.text = element_text(size = 18), legend.title =
                           element_text(size = 18), plot.title = element_text(size = 22)) +
  labs(title="Principal Component Analysis", subtitle = "Secretory Pathway") +
  scale_shape_manual(values=c(19,15,17,18)) +
  theme(plot.subtitle = element_text(size=15, face = "italic"))
dev.off()

```

PCA using met CellFie only
```{r}
### met Cellfie
cellfie.matrix.met <- filter(met.cellfie,`CHO-Core-1` >0) %>%
  as.matrix(rownames="Task")
PCA.cellfie.met <- PCA.plots(cellfie.matrix.met, pheno=pheno)
```

Plot loadings of PC1 
```{r}
## secretory scores only
PC1.sec <- as.data.frame(PCA.cellfie.sec$results$rotation) %>%
  tibble::rownames_to_column(var="Task") %>% select(c("Task", "PC1")) %>%
  arrange(PC1)
level.order <- arrange(PC1.sec, PC1)
PC1.sec <- mutate(PC1.sec, Task=factor(Task, levels=level.order$Task))

png("Figures/PC1sec_loadings.png", width=580, height=480)
ggplot(PC1.sec, aes(PC1, Task, fill=PC1)) +
  geom_bar(stat="identity") +
  ggtitle("PC1 Loadings") +
  scale_fill_gradient2(low=scales::muted("red"), mid="white", high=scales::muted("green"), name="Loading") +
  theme_linedraw() +
  theme(text=element_text(size=19), axis.title.y=element_blank(), axis.title.x=element_blank())
dev.off()

## met and sec scores
PC1 <- as.data.frame(PCA.cellfie$results$rotation) %>%
  tibble::rownames_to_column(var="Task") %>% select(c("Task", "PC1")) %>%
  arrange(PC1)
PC1.subset <- rbind(PC1[1:10,], PC1[(204-9):204,])
level.order <- arrange(PC1.subset, PC1)
PC1.subset <- mutate(PC1.subset, Task=factor(Task, levels=level.order$Task))

```

Plot loadings of PC2 
```{r}
## met and sec scores
PC2 <- as.data.frame(PCA.cellfie$results$rotation) %>%
  tibble::rownames_to_column(var="Task") %>% select(c("Task", "PC2")) %>%
  arrange(PC2)
PC2.subset <- rbind(PC2[1:10,], PC2[(204-9):204,])
level.order <- arrange(PC2.subset, PC2)
PC2.subset <- mutate(PC2.subset, Task=factor(Task, levels=level.order$Task))

## secretory scores only
PC2.sec <- as.data.frame(PCA.cellfie$results$rotation) %>%
  tibble::rownames_to_column(var="Task") %>% select(c("Task", "PC2")) %>%
  left_join(taskInfo) %>%
  filter(Pathway=="Secretory") %>%
  arrange(PC2)
level.order <- arrange(PC2.sec, PC2)
PC2.sec <- mutate(PC2.sec, Task=factor(Task, levels=level.order$Task))

png("Figures/PC2sec_loadings.png", width=640, height=480)
ggplot(PC2.sec, aes(PC2, Task, fill=PC2)) +
  geom_bar(stat="identity") +
  ggtitle("PC2 Secretory Task Loadings") +
  scale_fill_gradient2(low=scales::muted("purple"), mid="white", high=scales::muted("green"), name="Loading") +
  theme_linedraw() +
  theme(text=element_text(size=19), axis.title.y=element_blank(), axis.title.x=element_blank())
dev.off()
```

 
Clevland plot to show loadings of all secretory tasks on both PC1 and PC2
```{r}
PC1.sec <- left_join(PC1, taskInfo) %>%
  filter(Pathway=="Secretory") %>% 
  rename(Loading=PC1) %>% mutate(PC="PC1")
PC2.sec <- left_join(PC2, taskInfo) %>%
  filter(Pathway=="Secretory") %>% 
  rename(Loading=PC2) %>% mutate(PC="PC2")

#level.order <- arrange(DE.sec$EXPvsSTA_A1$welch, LFC) %>% na.omit()
PC1_PC2 <- rbind(PC1.sec, PC2.sec) #%>% mutate(Task=factor(Task, levels=level.order$Task))

clev.plot <- ggplot(PC1_PC2, aes(Loading, Task)) +
  geom_point(aes(color=PC), size=4, alpha=0.7) +
  theme(text=element_text(size=19), axis.title.y=element_blank(), legend.position = "bottom") +
  scale_color_manual(values=c("#E69F00", "#56B4E9")) +
  ggtitle("PC Loadings")


png("Figures/clevland_PC_grid.png", width=610, height=610)
clev.plot + theme(legend.position = "right") +
  geom_vline(xintercept = 0, linetype="dotted", color = "red", size=1) +
  facet_grid(System ~ ., scales="free", 
                       labeller = labeller(System = label_wrap_gen(width = 13)))                              # Change strip.background color
  #theme(strip.background = element_rect(fill = c("red", "blue", "yellow", "green", "pink")))
dev.off()
```
 
 
############################################################################
Differentially active tasks
############################################################################

Function
```{r}
welch.test <- function(group1, group2, cellfie, taskInfo, p.thresh, LFC.thresh){
  in.group1 <- select(cellfie, group1) %>% log2()
  in.group2 <- select(cellfie, group2) %>% log2()
  welch.df <- matrixTests::row_t_welch(in.group1, in.group2) %>% 
    data.table() %>%
    cbind(taskInfo) %>% 
    mutate(p.adjust := p.adjust(pvalue, method="BH")) %>% 
    mutate(LFC := mean.x-mean.y) #%>% mutate(LFC := log2(FC))
  
  sig.tasks <- select(welch.df, c("mean.x", "mean.y", "mean.diff","LFC", "p.adjust", names(taskInfo))) %>%
    filter(p.adjust <= p.thresh) %>% filter(LFC <= -log2(LFC.thresh) | LFC >= log2(LFC.thresh))
  
  return(list(welch=welch.df, sigRes=sig.tasks))
}
```

```{r}
# select groups for comparison: "A0_EXP" "A0_STA" "A1_EXP" "A1_STA"
#A1 (high producers)
#A0 (low producers)
group.A1_EXP <- pheno$sample.id[pheno$condition_state == "A1_EXP"] 
group.A1_STA <- pheno$sample.id[pheno$condition_state == "A1_STA"] 
group.A0_EXP <- pheno$sample.id[pheno$condition_state == "A0_EXP"] 
group.A0_STA <- pheno$sample.id[pheno$condition_state == "A0_STA"]  

## sec cellfie
sec.taskInfo <- data.frame(Task=sec.cellfie$Task) %>% left_join(taskInfo)
identical(sec.cellfie$Task, sec.taskInfo$Task)

DE.sec <- list(
  # high vs low in EXP
  A1vsA0_EXP=welch.test(group1=group.A1_EXP, group2=group.A0_EXP, cellfie=sec.cellfie, taskInfo=sec.taskInfo, p.thresh=0.1, LFC.thresh=1),
  # high vs low in STA
  A1vsA0_STA=welch.test(group1=group.A1_STA, group2=group.A0_STA, cellfie=sec.cellfie, taskInfo=sec.taskInfo, p.thresh=0.1, LFC.thresh=1),
  #EXP vs STA high producing cell line
  STAvsEXP_A1=welch.test(group1=group.A1_STA, group2=group.A1_EXP, cellfie=sec.cellfie, taskInfo=sec.taskInfo, p.thresh=0.1, LFC.thresh=1),
  #EXP vs STA high producing cell line
  STAvsEXP_A0=welch.test(group1=group.A0_STA, group2=group.A0_EXP, cellfie=sec.cellfie, taskInfo=sec.taskInfo, p.thresh=0.1, LFC.thresh=1)
  )
openxlsx::write.xlsx(list(A1vsA0_EXP=DE.sec$A1vsA0_EXP$sigRes,
                          A1vsA0_STA=DE.sec$A1vsA0_STA$sigRes,
                          STAvsEXP_A1=DE.sec$STAvsEXP_A1$sigRes,
                          STAvsEXP_A0=DE.sec$STAvsEXP_A0$sigRes), "Output/DEsig_sec.xlsx")
openxlsx::write.xlsx(list(A1vsA0_EXP=DE.sec$A1vsA0_EXP$welch,
                          A1vsA0_STA=DE.sec$A1vsA0_STA$welch,
                          STAvsEXP_A1=DE.sec$STAvsEXP_A1$welch,
                          STAvsEXP_A0=DE.sec$STAvsEXP_A0$welch), "Output/DE_sec.xlsx")

## met cellfiesec.taskInfo <- data.frame(Task=sec.cellfie$Task) %>% left_join(taskInfo)
met.taskInfo <- data.frame(Task=met.cellfie$Task) %>% left_join(taskInfo)
identical(met.cellfie$Task, met.taskInfo$Task)

DE.met <- list(
  # high vs low in EXP
  A1vsA0_EXP=welch.test(group1=group.A1_EXP, group2=group.A0_EXP, cellfie=met.cellfie, taskInfo=met.taskInfo, p.thresh=0.1, LFC.thresh=1.5),
  # high vs low in STA
  A1vsA0_STA=welch.test(group1=group.A1_STA, group2=group.A0_STA, cellfie=met.cellfie, taskInfo=met.taskInfo, p.thresh=0.1, LFC.thresh=1.5),
  #EXP vs STA high producing cell line
  STAvsEXP_A1=welch.test(group1=group.A1_STA, group2=group.A1_EXP, cellfie=met.cellfie, taskInfo=met.taskInfo, p.thresh=0.1, LFC.thresh=1.5),
  #EXP vs STA high producing cell line
  STAvsEXP_A0=welch.test(group1=group.A0_STA, group2=group.A0_EXP, cellfie=met.cellfie, taskInfo=met.taskInfo, p.thresh=0.1, LFC.thresh=1.5)
  )
openxlsx::write.xlsx(list(A1vsA0_EXP=DE.met$A1vsA0_EXP$sigRes,
                          A1vsA0_STA=DE.met$A1vsA0_STA$sigRes,
                          STAvsEXP_A1=DE.met$STAvsEXP_A1$sigRes,
                          STAvsEXP_A0=DE.met$STAvsEXP_A0$sigRes), "Output/DEsig_met.xlsx")
```

Dotplot of sig tasks
Secretory
```{r}
## EXP vs STA
STAvsEXP <- mutate(DE.sec$STAvsEXP_A1$sigRes, clone="A1") %>% 
  rbind(mutate(DE.sec$STAvsEXP_A0$sigRes, clone="A0")) %>%
  arrange(LFC) 
level.order <- arrange(STAvsEXP, LFC)
STAvsEXP <- mutate(STAvsEXP, Task=factor(Task, levels=level.order$Task))

p1 <- ggplot(STAvsEXP, aes(LFC, Task)) +
  geom_point(aes(color=clone), size=5, alpha=0.7) +
  theme(text=element_text(size=19), axis.title.y=element_blank()) +
  ggtitle("STA vs EXP") +
  geom_vline(xintercept = 0, linetype="dotted", color = "red", size=1)


## A1 vs A0
A1vsA0 <- mutate(DE.sec$A1vsA0_EXP$sigRes, phase="EXP") %>% 
  rbind(mutate(DE.sec$A1vsA0_STA$sigRes, phase="STA")) %>%
  arrange(LFC)
level.order <- arrange(A1vsA0, LFC)
A1vsA0 <- mutate(A1vsA0, Task=factor(Task, levels=level.order$Task))

p2 <- ggplot(A1vsA0, aes(LFC, Task)) +
  geom_point(aes(color=phase), size=5, alpha=0.7) +
  theme(text=element_text(size=19), axis.title.y=element_blank()) +
  ggtitle("A1 vs A0") + 
  scale_color_brewer(palette="Accent") +
  geom_vline(xintercept = 0, linetype="dotted", color = "red", size=1)
```

Metabolic
```{r}
## EXP vs STA
STAvsEXP <- mutate(DE.met$STAvsEXP_A1$sigRes, clone="A1") %>% 
  rbind(mutate(DE.met$STAvsEXP_A0$sigRes, clone="A0")) %>%
  arrange(LFC) 
level.order <- arrange(STAvsEXP, LFC)
STAvsEXP <- mutate(STAvsEXP, Task=factor(Task, levels=level.order$Task))

p3 <- ggplot(STAvsEXP, aes(LFC, Task)) +
  geom_point(aes(color=clone), size=5, alpha=0.7) +
  theme(text=element_text(size=19), axis.title.y=element_blank()) +
  ggtitle("STA vs EXP") +
  geom_vline(xintercept = 0, linetype="dotted", color = "red", size=1)

## A1 vs A0
A1vsA0 <- mutate(DE.met$A1vsA0_EXP$sigRes, phase="EXP") %>% 
  rbind(mutate(DE.met$A1vsA0_STA$sigRes, phase="STA")) %>%
  arrange(LFC)
level.order <- arrange(A1vsA0, LFC)
A1vsA0 <- mutate(A1vsA0, Task=factor(Task, levels=level.order$Task))

p4 <- ggplot(A1vsA0, aes(LFC, Task)) +
  geom_point(aes(color=phase), size=5, alpha=0.7) +
  theme(text=element_text(size=19), axis.title.y=element_blank()) +
  ggtitle("A1 vs A0") + 
  scale_color_brewer(palette="Accent") +
  geom_vline(xintercept = 0, linetype="dotted", color = "red", size=1)

```

Combine all plots into a single figure
```{r}
p1.grob <- p1 + theme(plot.title = element_text(hjust = 0.5))#+ggtitle("")
p2.grob <- p2 + theme(plot.title = element_text(hjust = 0.5))#+ggtitle("")
p3.grob <- p3 + theme(plot.title = element_text(hjust = 0.5))#+ggtitle("")
p4.grob <- p4 + theme(plot.title = element_text(hjust = 0.5))#+ggtitle("")

png("Figures/dotplot_all.png", width=1200, height=700)
egg::ggarrange(p1.grob, p2.grob, p3.grob, p4.grob, ncol=2)
dev.off()
```

```{r}
STAvsEXP.sec <- mutate(DE.sec$STAvsEXP_A1$sigRes, condition="A1") %>% 
  rbind(mutate(DE.sec$STAvsEXP_A0$sigRes, condition="A0")) %>%
  mutate(comparison="STA vs. EXP") %>%
  arrange(LFC) 
A1vsA0.sec <- mutate(DE.sec$A1vsA0_EXP$sigRes, condition="EXP") %>% 
  rbind(mutate(DE.sec$A1vsA0_STA$sigRes, condition="STA")) %>%
  mutate(comparison="A1 vs. A0") %>%
  arrange(LFC)
STAvsEXP.met <- mutate(DE.met$STAvsEXP_A1$sigRes, condition="A1") %>% 
  rbind(mutate(DE.met$STAvsEXP_A0$sigRes, condition="A0")) %>%
  mutate(comparison="STA vs. EXP") %>%
  arrange(LFC)
A1vsA0.met <- mutate(DE.met$A1vsA0_EXP$sigRes, condition="EXP") %>% 
  rbind(mutate(DE.met$A1vsA0_STA$sigRes, condition="STA")) %>%
  mutate(comparison="A1 vs. A0") %>%
  arrange(LFC)

compare.all <- rbind(A1vsA0.met, A1vsA0.sec, STAvsEXP.met, STAvsEXP.sec) %>%
  mutate(condition=as.factor(condition))
levels(compare.all$condition) <- c(levels(compare.all$condition), "A1")

png("Figures/dotplotAll_grid.png", width=1100, height=550)
ggplot(compare.all, aes(LFC, Task)) +
  geom_point(aes(color=condition), size=6) +
  theme(text=element_text(size=21), axis.title.y=element_blank()) +
  scale_color_manual(values=c("#7FC97F", "#BEAED4", "#FDC086", "#386CB0") %>%
                       setNames(levels(compare.all$condition)), drop=F, 
                     breaks=c('A1', 'A0', 'EXP', 'STA')) +
  #scale_color_brewer(palette="Accent") +
  geom_vline(xintercept = 0, linetype="dotted", color = "red", size=1) +
  facet_grid(Pathway ~ comparison, margins=F, scales="free", space="free_y")
dev.off()

```


############################################################################
DE A1 vs A0
############################################################################

```{r}
library(DESeq2)
```

Load data
```{r}
raw.counts <- fread("Input/counts.csv") %>% select(c("gene.id", "gene", colnames(count.matrix)))
pheno <- openxlsx::read.xlsx("Input/Susan_Annotations.xlsx", sheet="clean")

# It is absolutely critical that the columns of the count matrix and the rows of the column data (information about samples) are in the same order.
identical(colnames(count.matrix), pheno$sample.id)
all(pheno$sample.id == colnames(count.matrix))
```

Construct DESeqDataSet from count matrix
```{r}
dds <- DESeqDataSetFromMatrix(countData = count.matrix,
                              colData = pheno,
                              design = ~condition)
```

Pre-filter and set condition level
```{r}
keep <- rowSums(counts(dds)) >=10
dds <- dds[keep,]
dds$condition <- relevel(dds$condition, ref = "A0")
```

Differential expression analysis
```{r}
set.seed(18)
dds <- DESeq(dds)
res <- results(dds) %>% as.data.frame() %>% tibble::rownames_to_column(var="gene")
res.sig <- filter(res, padj<= 0.01)
```

Add PSIM info
```{r}
CHOmap <- openxlsx::read.xlsx("Input/CHOProtein.xlsx")
CHOmap2 <- fread("Input/geneInfo_cho_simple.tsv") %>%
  mutate(GeneID=as.character(GeneID))
PSIM <- openxlsx::read.xlsx("Input/PSIM_matrix.xlsx")

res.PSIM <- left_join(res, CHOmap, by=c("gene"="ENTREZ_GENE")) %>%
  left_join(PSIM, by=c("UNIPROTKB"="Entry")) %>% filter(GPI==1) %>% arrange(log2FoldChange) %>% 
  left_join(CHOmap2, by=c("gene"="GeneID"))
View(filter(res.PSIM, padj<=0.01) %>% filter(abs(log2FoldChange) >= log2(1.5)))

```
 
Heatmap bar of GPI proteins 
```{r}
#my_palette <- colorRampPalette(c(scales::muted("red"), "white", scales::muted("green")))(n = 100)
my_palette <- colorRampPalette(c(scales::muted("purple"), "white", scales::muted("green")))(n = 100)

png("Figures/GPI_long.png", width=300, height=1200)
gplots::heatmap.2(cbind(res.PSIM$log2FoldChange, res.PSIM$log2FoldChange), trace="n", Colv=NA, 
                  labRow = res.PSIM$Symbol, dendrogram = "none", col=my_palette, margins=c(12,15),
                  labCol="", srtRow=-45, cexRow = 3, key=F, Rowv = F)#, sepcolor = "black", rowsep=c(0:20), colsep=c(0,2))
dev.off()

png("Figures/GPI_key.png", width=900, height=700)
gplots::heatmap.2(cbind(res.PSIM$log2FoldChange, res.PSIM$log2FoldChange), trace="n", Colv=NA, 
                  labRow = res.PSIM$Symbol, dendrogram = "none", col=my_palette,margins=c(6,8),
                  labCol="", srtRow=-45, cexRow = 1.5, key.title = "LFC", keysize = 4, density.info = "density",
                  key.par=list(cex=1.5), denscol="black", densadj = 0.8, Rowv = F)
dev.off()
```

