---
title: "R Notebook"
output: html_notebook
---

```{r}
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE)
```

```{r}
library(data.table)
library(dplyr)
library(factoextra)
library(ggplot2)
library(gplots)
library(pheatmap)
library(RColorBrewer)
```

Load original TPM data (for labels) and pheno data
```{r}
#TPM <- fread("~/Desktop/SenateExam/Aim2_CellFie/ImmCellFie_Manuscript/Output/demo_expression.csv")
TPM <- fread("Output/HPAblood_nTPMexpression.csv")
pheno <- fread("~/Desktop/SenateExam/Aim2_CellFie/ImmCellFie_Manuscript/Output/demo_pheno_cellType.csv", colClasses=c("character")) %>%
  dplyr::select(c("Immune cell", "Cell_Type")) %>% unique() %>% rename("Immune_Cell"="Immune cell") %>% 
  filter(Immune_Cell!="total PBMC")

identical(colnames(TPM)[-1], pheno$Immune_Cell)

```

Load Cellfie results
```{r}
## Secretory Pathway
#sec.cellfie <- fread("~/Desktop/SenateExam/Aim2_CellFie/ImmCellFie_Manuscript/Output/secretoryCellFie/mean/score.csv") %>%
sec.cellfie <- fread("CellFie_output/blood/secretoryPathway/mean/TPM/score.csv") %>%
  setNames(colnames(TPM)[-1])
sec.cellfie.binary <- fread("CellFie_output/blood/secretoryPathway/mean/TPM/score_binary.csv") %>%
  setNames(colnames(TPM)[-1])
sec.taskInfo <-  openxlsx::read.xlsx("/Users/helenmasson/Desktop/SenateExam/Aim2_CellFie/SecPathway.CellFIE/usingGeM/Generate_essentialRxns_manual/essentialRxn_notebook2022.xlsx", sheet="TaskInfo") %>% mutate(Pathway="Secretory Pathway")
sec.cellfie.filter <- sec.cellfie*sec.cellfie.binary

## Metabolic Pathway
#met.cellfie <- fread("~/Desktop/SenateExam/Aim2_CellFie/ImmCellFie_Manuscript/Output/metCellFie/mean/score.csv") %>%
met.cellfie <- fread("CellFie_output/blood/metabolicPathway/mean/TPM/score.csv") %>%
  setNames(colnames(TPM)[-1])
met.cellfie.binary <- fread("CellFie_output/blood/metabolicPathway/mean/TPM/score_binary.csv") %>%
  setNames(colnames(TPM)[-1])
met.taskInfo <- fread("CellFie_output/blood/metabolicPathway/mean/TPM/taskInfo.csv") %>%
  mutate(Pathway="Metabolic Pathway") %>%
  setNames(colnames(sec.taskInfo))
met.cellfie.filter <- met.cellfie*met.cellfie.binary

## Combine
res.cellfie <- rbind(met.cellfie, sec.cellfie)
res.cellfie.binary <- rbind(met.cellfie.binary, sec.cellfie.binary)
res.cellfie.filter <- res.cellfie * res.cellfie.binary
taskInfo <- rbind(met.taskInfo, sec.taskInfo)
```

remove total PBMC
```{r}
#pheno <- pheno[Cell_Type != "total PBMC" ,]
#TPM <- dplyr::select(TPM, c("Gene", pheno$`Sample ID`))

#res.cellfie <- dplyr::select(res.cellfie, pheno$`Sample ID`)
#res.cellfie.binary <- dplyr::select(res.cellfie.binary, pheno$`Sample ID`)
#res.cellfie.filter <- dplyr::select(res.cellfie.filter, pheno$`Sample ID`)

#sec.cellfie <- dplyr::select(sec.cellfie, pheno$`Sample ID`)
#sec.cellfie.binary <- dplyr::select(sec.cellfie.binary, pheno$`Sample ID`)
#sec.cellfie.filter <- dplyr::select(sec.cellfie.filter, pheno$`Sample ID`)
```


####################################################################
Principal Component Analysis
####################################################################

PCA using TPM
```{r}
### prepare expression data for PCA
TPM.matrix <- as.matrix(TPM[, -1]) %>% `rownames<-`(TPM$Gene) %>% `colnames<-`(pheno$Immune_Cell)
# remove constant/zero columns
TPM.matrix <- TPM.matrix[matrixStats::rowSds(TPM.matrix) > 0.00001 ,]

### define cell line labels and colors
#cellLine.labels <- as.factor(c(rep("NAFLD",3), rep("WT", 3)))
#cellLine_col <- brewer.pal(4,"Dark2")[as.numeric(cellLine.labels)]

# PCA
res.pca <- prcomp(t(TPM.matrix), scale=T)

# visualize eigenvalues/variances
fviz_screeplot(res.pca, addlabels=T, ylim = c(0, 100))

# plot individuals on first two components
fviz_pca_ind(res.pca,
             geom.ind = c("text", "point"),
             pointsize = 3,
             mean.point = FALSE,
             show.legend=FALSE,
             habillage = as.factor(pheno$Immune_Cell),
             #palette = "Dark2",
             legend.title = "Donor",
             repel = TRUE     # Avoid text overlapping
) + theme(text=element_text(size=4), axis.text = element_text(size = 10), axis.title = element_text(size = 10), legend.text = element_text(size = 10), legend.title = element_text(size = 10), plot.title = element_text(size = 20))

## Variable contributions to the principal axes
# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 5)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 5)
```

PCA using CellFie
```{r}
### prepare expression data for PCA
cellfie.matrix <- as.matrix(res.cellfie) %>% `rownames<-`(taskInfo$Task) %>% `colnames<-`(pheno$Immune_Cell) %>%
  na.omit()
#cellfie.matrix <- as.matrix(res.cellfie.filter) %>% `rownames<-`(taskInfo$Task) %>% `colnames<-`(pheno$unique_Imm)

# remove constant/zero columns
cellfie.matrix <- cellfie.matrix[matrixStats::rowSds(cellfie.matrix) > 0.00001 ,]

# PCA
res.pca <- prcomp(t(cellfie.matrix), scale=T)

# visualize eigenvalues/variances
fviz_screeplot(res.pca, addlabels=T, ylim = c(0, 100))

# plot individuals on first two components
fviz_pca_ind(res.pca,
             geom.ind = c("text", "point"),
             pointsize = 3,
             mean.point = FALSE,
             show.legend=FALSE,
             habillage = as.factor(pheno$Immune_Cell),
             #palette = "Dark2",
             legend.title = "Cell Line",
             repel = TRUE     # Avoid text overlapping
) + theme(text=element_text(size=4), axis.text = element_text(size = 10), axis.title = element_text(size = 10), legend.text = element_text(size = 10), legend.title = element_text(size = 10), plot.title = element_text(size = 20))

## Variable contributions to the principal axes
# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 5)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 5)
```


####################################################################
Kmeans Clustering
####################################################################

Kmeans on Immune cell type

Using TPM
```{r}
## Determining optimal number of clusters
TPM.scale <- scale(t(TPM.matrix))
# elbow method
fviz_nbclust(TPM.scale, kmeans, method = "wss", k.max = 15)
# silhouette method
fviz_nbclust(TPM.scale, kmeans, method = "silhouette", k.max = 15)
# gap statistic
#fviz_gap_stat(cluster::clusGap(TPM.scale, FUN=kmeans, nstart=25, K.max = 15, B=10))
```
Use optimal k
```{r}
set.seed(17)
kmeans.TPM.type <- kmeans(TPM.scale, 6, nstart = 25)
```


Using CellFie
```{r}
## Determining optimal number of clusters
cellfie.scale <- scale(t(cellfie.matrix))
# elbow method
fviz_nbclust(cellfie.scale, kmeans, method = "wss", k.max = 15)
# silhouette method
fviz_nbclust(cellfie.scale, kmeans, method = "silhouette", k.max = 15)
# gap statistic
#fviz_gap_stat(cluster::clusGap(cellfie.scale, FUN=kmeans, nstart=25, K.max = 15, B=50))
```
Use optimalk
```{r}
set.seed(18)
kmeans.cellfie.type <- kmeans(cellfie.scale, 6, nstart = 25)
```

combine all clustering methods into a single dataframe
```{r}
cluster.TPM.type <- as.data.frame(kmeans.TPM.type$cluster) %>% 
  tibble::rownames_to_column(var="Immune_Cell") %>%
  dplyr::rename("TPM.type"="kmeans.TPM.type$cluster") %>%
  mutate(TPM.type=as.factor(TPM.type))
  
cluster.cellfie.type <- as.data.frame(kmeans.cellfie.type$cluster) %>% 
  tibble::rownames_to_column(var="Immune_Cell") %>%
  dplyr::rename("cellfie.type"="kmeans.cellfie.type$cluster") %>%
  mutate(cellfie.type=as.factor(cellfie.type))

pheno.clusters <- left_join(pheno, cluster.TPM.type) %>% left_join(cluster.cellfie.type) %>%
  arrange(Cell_Type) 
```

bar plot cluster by immune cell
```{r}
# TPM
ggplot(data = pheno.clusters, aes(y = TPM.type)) +
  geom_bar(aes(fill =Cell_Type)) +
  ggtitle("Clusters by Immune Cell Type using TPM") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(text=element_text(size=15))

# cellfie
ggplot(data = pheno.clusters, aes(y = cellfie.type)) +
  geom_bar(aes(fill =Cell_Type)) +
  ggtitle("Clusters by Immune Cell Type using CellFie") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(text=element_text(size=15))
```
 
 bar plot immune cell by cluster
```{r}
# TPM
ggplot(data = pheno.clusters, aes(y =Cell_Type)) +
  geom_bar(aes(fill = TPM.cluster)) +
  ggtitle("Clusters by Immune Cell Type using TPM") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(text=element_text(size=15))

# cellfie
ggplot(data = pheno.clusters, aes(y =Cell_Type)) +
  geom_bar(aes(fill = cellfie.cluster)) +
  ggtitle("Clusters by Immune Cell Type using CellFie") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(text=element_text(size=15))
```

Clustering on subset of cells
```{r}
bob <- filter(pheno, Cell_Type == "Granulocytes" | Cell_Type == "Dentritic Cells" | Cell_Type == "NK-cells")
sam <- cellfie.matrix[, match(bob$unique_Imm, colnames(cellfie.matrix))]
identical(bob$unique_Imm, colnames(sam))

#remove rows/tasks with no variation
sam <- sam[matrixStats::rowSds(sam) > 0.000000001 ,]

set.seed(182)
kmeans.subset <- kmeans(scale(t(sam)), length(unique(bob$`Immune cell`)), nstart = 25)

cluster.subset <- as.data.frame(kmeans.subset$cluster) %>% 
  tibble::rownames_to_column(var="unique_Imm") %>%
  dplyr::rename("cluster"="kmeans.subset$cluster") %>%
  mutate(cluster=as.factor(cluster)) %>% right_join(bob)

ggplot(data = cluster.subset, aes(y = cluster)) +
  geom_bar(aes(fill = `Immune cell`)) +
  ggtitle("Clusters by Immune Cell Type using CellFie") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(text=element_text(size=15))
```



####################################################################
Hierarchical Clustering
####################################################################

```{r}
#using all cellfie
set.seed(18)
hclust.cellfie <- pheatmap::pheatmap(cellfie.matrix, 
                   main="Hierarchical clustering using CellFie", 
                   scale="row",
                   border_color = NA,
                   #cluster.cols = hclust(dist(cellfie.matrix, method="euclidean"), method="complete"),
                   clustering_distance_cols = "correlation",
                   clustering_method = "complete",
                   annotation_col = data.frame(cell=pheno$`Immune cell`, cellType=pheno$Cell_Type) %>% `rownames<-`(pheno$unique_Imm),
                   show_colnames = F,
                   show_rownames = F,
                   treeheight_row = 0,
                   legend_breaks = c(0,1))

#using met cellfie only
set.seed(18)
hclust.metCellfie <- pheatmap::pheatmap(cellfie.matrix[rownames(cellfie.matrix) %in% met.taskInfo$Task ,], 
                   main="Hierarchical clustering using metabolic CellFie only", 
                   scale="row",
                   border_color = NA,
                   #cluster.cols = hclust(dist(cellfie.matrix, method="euclidean"), method="complete"),
                   clustering_distance_cols = "correlation",
                   clustering_method = "complete",
                   annotation_col = data.frame(cell=pheno$`Immune cell`, cellType=pheno$Cell_Type) %>% `rownames<-`(pheno$unique_Imm),
                   show_colnames = F,
                   show_rownames = F,
                   treeheight_row = 0,
                   legend_breaks = c(0,1))

#using TPM
set.seed(18)
hclust.TPM <- pheatmap::pheatmap(TPM.matrix, 
                   main="Hierarchical clustering using TPM", 
                   scale="row",
                   border_color = NA,
                   clustering_distance_cols = "correlation",
                   clustering_method = "complete",
                   annotation_col = data.frame(cell=pheno$`Immune cell`, cellType=pheno$Cell_Type) %>% `rownames<-`(pheno$unique_Imm),
                   show_colnames = F,
                   show_rownames = F,
                   treeheight_row = 0,
                   legend_breaks = c(0,1))
```

subset of cells
```{r}
# using sec and met tasks
set.seed(18)
hclust.cellfie <- pheatmap::pheatmap(sam, 
                   main="Hierarchical clustering using CellFie", 
                   color = hcl.colors(100, "BluYl"),
                   scale="row", # "none" or "row"
                   border_color = NA,
                   #cluster.cols = hclust(dist(cellfie.matrix, method="euclidean"), method="complete"),
                   clustering_distance_cols = "correlation",
                   clustering_method = "complete",
                   annotation_col = data.frame(cell=bob$`Immune cell`, cellType=bob$Cell_Type) %>%
                     `rownames<-`(bob$unique_Imm),
                   #annotation_row = tibble::column_to_rownames(taskInfo, var="Task") %>% select(Pathway),
                   show_colnames = F,
                   show_rownames = F,
                   treeheight_row = 0,
                   legend_breaks = c(0,1))

#using met tasks
set.seed(18)
hclust.cellfie <- pheatmap::pheatmap(sam[rownames(sam) %in% met.taskInfo$Task ,], 
                   main="Hierarchical clustering using metabolic CellFie", 
                   color = hcl.colors(100, "BluYl"),
                   scale="row", # "none" or "row"
                   border_color = NA,
                   #cluster.cols = hclust(dist(cellfie.matrix, method="euclidean"), method="complete"),
                   clustering_distance_cols = "correlation",
                   clustering_method = "complete",
                   annotation_col = data.frame(cell=bob$`Immune cell`, cellType=bob$Cell_Type) %>%
                     `rownames<-`(bob$unique_Imm),
                   #annotation_row = tibble::column_to_rownames(taskInfo, var="Task") %>% select(Pathway),
                   show_colnames = F,
                   show_rownames = F,
                   treeheight_row = 0,
                   legend_breaks = c(0,1))

#using sec tasks
set.seed(18)
hclust.cellfie <- pheatmap::pheatmap(sam[rownames(sam) %in% sec.taskInfo$Task ,], 
                   main="Hierarchical clustering using secretory CellFie", 
                   color = hcl.colors(100, "BluYl"),
                   scale="row", # "none" or "row"
                   border_color = NA,
                   #cluster.cols = hclust(dist(cellfie.matrix, method="euclidean"), method="complete"),
                   clustering_distance_cols = "correlation",
                   clustering_method = "complete",
                   annotation_col = data.frame(cell=bob$`Immune cell`, cellType=bob$Cell_Type) %>%
                     `rownames<-`(bob$unique_Imm),
                   #annotation_row = tibble::column_to_rownames(taskInfo, var="Task") %>% select(Pathway),
                   show_colnames = F,
                   show_rownames = F,
                   treeheight_row = 0,
                   legend_breaks = c(0,1))

```

