---
title: "R Notebook"
output: html_notebook
---

```{r}
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE)
```

```{r}
library(data.table)
library(dplyr)
library(factoextra)
library(ggplot2)
library(gplots)
library(pheatmap)
library(RColorBrewer)
```

Load original TPM data (for labels) and pheno data
```{r}
#TPM <- fread("~/Desktop/SenateExam/Aim2_CellFie/ImmCellFie_Manuscript/Output/demo_expression.csv")
TPM <- fread("Output/HPAblood_109_TPMexpression.csv")
pheno <- fread("Output/HPAblood_109_pheno_withType.csv", colClasses=c("character"))

identical(colnames(TPM)[-1], pheno$Sample_ID)
pheno <- pheno[match(colnames(TPM)[-1], pheno$Sample_ID) ,]
identical(colnames(TPM)[-1], pheno$Sample_ID)
```

Load Cellfie results
```{r}
## Secretory Pathway
#sec.cellfie <- fread("~/Desktop/SenateExam/Aim2_CellFie/ImmCellFie_Manuscript/Output/secretoryCellFie/mean/score.csv") %>%
sec.cellfie <- fread("CellFie_output/blood/secretoryPathway/mean/109_TPM/score.csv") %>%
  setNames(colnames(TPM)[-1])
sec.cellfie.binary <- fread("CellFie_output/blood/secretoryPathway/mean/109_TPM/score_binary.csv") %>%
  setNames(colnames(TPM)[-1])
sec.taskInfo <-  openxlsx::read.xlsx("/Users/helenmasson/Desktop/SenateExam/Aim2_CellFie/SecPathway.CellFIE/usingGeM/Generate_essentialRxns_manual/essentialRxn_notebook2022.xlsx", sheet="TaskInfo") %>% mutate(Pathway="Secretory Pathway")
sec.cellfie.filter <- sec.cellfie*sec.cellfie.binary

## Metabolic Pathway
#met.cellfie <- fread("~/Desktop/SenateExam/Aim2_CellFie/ImmCellFie_Manuscript/Output/metCellFie/mean/score.csv") %>%
met.cellfie <- fread("CellFie_output/blood/metabolicPathway/mean/109_TPM/score.csv") %>%
  setNames(colnames(TPM)[-1])
met.cellfie.binary <- fread("CellFie_output/blood/metabolicPathway/mean/109_TPM/score_binary.csv") %>%
  setNames(colnames(TPM)[-1])
met.taskInfo <- fread("CellFie_output/blood/metabolicPathway/mean/109_TPM/taskInfo.csv") %>%
  mutate(Pathway="Metabolic Pathway") %>%
  setNames(colnames(sec.taskInfo))
met.cellfie.filter <- met.cellfie*met.cellfie.binary

## Combine
res.cellfie <- rbind(met.cellfie, sec.cellfie)
res.cellfie.binary <- rbind(met.cellfie.binary, sec.cellfie.binary)
res.cellfie.filter <- res.cellfie * res.cellfie.binary
taskInfo <- rbind(met.taskInfo, sec.taskInfo)
```

####################################################################
Principal Component Analysis
####################################################################

PCA using TPM
```{r}
### prepare expression data for PCA
TPM.matrix <- as.matrix(TPM[, -1]) %>% `rownames<-`(TPM$ENTREZID) %>% `colnames<-`(pheno$Sample_ID)
# remove constant/zero columns
TPM.matrix <- TPM.matrix[matrixStats::rowSds(TPM.matrix) > 0.001 ,]

# PCA
res.pca <- prcomp(t(TPM.matrix), scale=T)

# visualize eigenvalues/variances
fviz_screeplot(res.pca, addlabels=T, ylim = c(0, 100))

# plot individuals on first two components
fviz_pca_ind(res.pca,
             geom.ind = c("text", "point"),
             label="none",
             pointsize = 3,
             mean.point = FALSE,
             show.legend=FALSE,
             habillage = as.factor(pheno$Immune_Cell),
             #palette = "Dark2",
             legend.title = "Donor",
             repel = TRUE     # Avoid text overlapping
) + theme(text=element_text(size=4), axis.text = element_text(size = 10), axis.title = element_text(size = 10), legend.text = element_text(size = 10), legend.title = element_text(size = 10), plot.title = element_text(size = 20))

## Variable contributions to the principal axes
# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 5)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 5)
```

PCA using CellFie
```{r}
### prepare expression data for PCA
cellfie.matrix <- as.matrix(res.cellfie) %>% `rownames<-`(taskInfo$Task) %>% `colnames<-`(pheno$Sample_ID)

# remove constant/zero columns
cellfie.matrix <- cellfie.matrix[matrixStats::rowSds(cellfie.matrix) != 0 ,]

# PCA
res.pca <- prcomp(t(cellfie.matrix), scale=T)

# visualize eigenvalues/variances
fviz_screeplot(res.pca, addlabels=T, ylim = c(0, 100))

# plot individuals on first two components
fviz_pca_ind(res.pca,
             geom.ind = c("text", "point"),
             label="none",
             pointsize = 3,
             mean.point = FALSE,
             show.legend=FALSE,
             habillage = as.factor(pheno$Immune_Cell),
             #palette = "Dark2",
             legend.title = "Cell Line",
             repel = TRUE     # Avoid text overlapping
) + theme(text=element_text(size=4), axis.text = element_text(size = 10), axis.title = element_text(size = 10), legend.text = element_text(size = 10), legend.title = element_text(size = 10), plot.title = element_text(size = 20))

## Variable contributions to the principal axes
# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 5)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 5)
```

####################################################################
Hierarchical Clustering
####################################################################

It is very difficult to get 18 distinct colors
Manually attempt to get most distinct colors as possible
```{r}
library(randomcoloR)
palette <- distinctColorPalette(18)
#pie(rep(1, 18), col=palette)

ann_colors <- list(cell=c(basophil="#E3589A", `classical monocyte`="#DA4FDF", eosinophil="#D86049", `gdT-cell`="#764ADF", `intermediate monocyte`="#7AE94E", `MAIT T-cell`="#D3BEDC", `memory B-cell`="#7FBFDC", `memory CD4 T-cell`="#D291E2", `memory CD8 T-cell`="#7E5C99", `myeloid DC`="#E0989B", `naive B-cell`="#C4E090", `naive CD4 T-cell`="#DEAF66",
`naive CD8 T-cell`="#D8DF52", neutrophil="#DDE0CC", `NK-cell`="#7DE0D1", `non-classical monocyte`="#809173", `plasmacytoid DC`="#6CDE90", `T-reg`="#6E95DE"))
```


```{r}
#using all cellfie
png("Figures/hclust_CellFie.png", width=1200, height=900)
set.seed(18)
hclust.cellfie <- pheatmap::pheatmap(cellfie.matrix,
                    fontsize=20,
                    main="Hierarchical clustering using CellFie", 
                    scale="row",
                    border_color = NA,
                    #cluster.cols = hclust(dist(cellfie.matrix, method="euclidean"), method="complete"),
                    clustering_distance_cols = "correlation",
                    clustering_method = "complete",
                    annotation_col = data.frame(cell=factor(pheno$Immune_Cell), cellType=pheno$Cell_Type) %>% `rownames<-`(pheno$Sample_ID),
                    annotation_colors = ann_colors,
                    show_colnames = F,
                    show_rownames = F,
                    treeheight_row = 0,
                    legend_breaks = c(0,1))
dev.off()
```
```{r}
#using met cellfie only
png("Figures/hclust_metCellFie.png", width=1200, height=900)
set.seed(18)
hclust.metCellfie <- pheatmap::pheatmap(cellfie.matrix[rownames(cellfie.matrix) %in% met.taskInfo$Task ,],
                  fontsize=20,
                  main="Hierarchical clustering using metabolic CellFie only", 
                  scale="row",
                  border_color = NA,
                  #cluster.cols = hclust(dist(cellfie.matrix, method="euclidean"), method="complete"),
                  clustering_distance_cols = "correlation",
                  clustering_method = "complete",
                  annotation_col = data.frame(cell=pheno$Immune_Cell, cellType=pheno$Cell_Type) %>% `rownames<-`(pheno$Sample_ID),
                  annotation_colors = ann_colors,
                  show_colnames = F,
                  show_rownames = F,
                  treeheight_row = 0,
                  legend_breaks = c(0,1))
dev.off()
```
```{r}
#using TPM
png("Figures/hclust_TPM.png", width=1200, height=900)
set.seed(18)
hclust.TPM <- pheatmap::pheatmap(TPM.matrix, 
                    fontsize=20,
                    main="Hierarchical clustering using TPM", 
                    scale="row",
                    border_color = NA,
                    clustering_distance_cols = "correlation",
                    clustering_method = "complete",
                    annotation_col = data.frame(cell=pheno$Immune_Cell, cellType=pheno$Cell_Type) %>% `rownames<-`(pheno$Sample_ID),
                    annotation_colors = ann_colors,
                    show_colnames = F,
                    show_rownames = F,
                    treeheight_row = 0,
                    legend_breaks = c(0,1))

dev.off()
```
















subset of cells
```{r}
# using sec and met tasks
#png("Figures/Figure_5a.png", width=800, height=1500)
set.seed(18)
hclust.cellfie <- pheatmap::pheatmap(sam, 
                   main="Hierarchical clustering using CellFie", 
                   color = hcl.colors(100, "BluYl"),
                   scale="row", # "none" or "row"
                   border_color = NA,
                   #cluster.cols = hclust(dist(cellfie.matrix, method="euclidean"), method="complete"),
                   clustering_distance_cols = "correlation",
                   clustering_method = "complete",
                   annotation_col = data.frame(cell=bob$`Immune cell`, cellType=bob$Cell_Type) %>%
                     `rownames<-`(bob$unique_Imm),
                   #annotation_row = tibble::column_to_rownames(taskInfo, var="Task") %>% select(Pathway),
                   show_colnames = F,
                   show_rownames = F,
                   treeheight_row = 0,
                   legend_breaks = c(0,1))
#dev.off()

#using met tasks
#png("Figures/Figure_5a.png", width=800, height=1500)
set.seed(18)
hclust.cellfie <- pheatmap::pheatmap(sam[rownames(sam) %in% met.taskInfo$Task ,], 
                   main="Hierarchical clustering using metabolic CellFie", 
                   color = hcl.colors(100, "BluYl"),
                   scale="row", # "none" or "row"
                   border_color = NA,
                   #cluster.cols = hclust(dist(cellfie.matrix, method="euclidean"), method="complete"),
                   clustering_distance_cols = "correlation",
                   clustering_method = "complete",
                   annotation_col = data.frame(cell=bob$`Immune cell`, cellType=bob$Cell_Type) %>%
                     `rownames<-`(bob$unique_Imm),
                   #annotation_row = tibble::column_to_rownames(taskInfo, var="Task") %>% select(Pathway),
                   show_colnames = F,
                   show_rownames = F,
                   treeheight_row = 0,
                   legend_breaks = c(0,1))
#dev.off()

#using sec tasks
#png("Figures/Figure_5a.png", width=800, height=1500)
set.seed(18)
hclust.cellfie <- pheatmap::pheatmap(sam[rownames(sam) %in% sec.taskInfo$Task ,], 
                   main="Hierarchical clustering using secretory CellFie", 
                   color = hcl.colors(100, "BluYl"),
                   scale="row", # "none" or "row"
                   border_color = NA,
                   #cluster.cols = hclust(dist(cellfie.matrix, method="euclidean"), method="complete"),
                   clustering_distance_cols = "correlation",
                   clustering_method = "complete",
                   annotation_col = data.frame(cell=bob$`Immune cell`, cellType=bob$Cell_Type) %>%
                     `rownames<-`(bob$unique_Imm),
                   #annotation_row = tibble::column_to_rownames(taskInfo, var="Task") %>% select(Pathway),
                   show_colnames = F,
                   show_rownames = F,
                   treeheight_row = 0,
                   legend_breaks = c(0,1))
#dev.off()
```





####################################################################
Kmeans Clustering
####################################################################

Kmeans on "Immune Cell"
```{r}
set.seed(182)
kmeans.cellfie <- kmeans(scale(t(cellfie.matrix)), length(unique(pheno$Immune_Cell)), nstart = 25)
set.seed(17)
kmeans.TPM <- kmeans(scale(t(TPM.matrix)), length(unique(pheno$Immune_Cell)), nstart = 25)
```

Kmeans on Immune cell type
```{r}
set.seed(182)
kmeans.cellfie.type <- kmeans(scale(t(cellfie.matrix)), length(unique(pheno$Cell_Type)), nstart = 25)

set.seed(17)
kmeans.TPM.type <- kmeans(scale(t(TPM.matrix)), length(unique(pheno$Cell_Type)), nstart = 25)
```

combine all clustering methods into a single dataframe
```{r}
cluster.TPM <- as.data.frame(kmeans.TPM$cluster) %>% 
  tibble::rownames_to_column(var="Sample_ID") %>%
  dplyr::rename("TPM.cluster"="kmeans.TPM$cluster") %>%
  mutate(TPM.cluster=as.factor(TPM.cluster))

cluster.TPM.type <- as.data.frame(kmeans.TPM.type$cluster) %>% 
  tibble::rownames_to_column(var="Sample_ID") %>%
  dplyr::rename("TPM.type"="kmeans.TPM.type$cluster") %>%
  mutate(TPM.type=as.factor(TPM.type))
  
cluster.cellfie <- as.data.frame(kmeans.cellfie$cluster) %>% 
  tibble::rownames_to_column(var="Sample_ID") %>%
  dplyr::rename("cellfie.cluster"="kmeans.cellfie$cluster") %>%
  mutate(cellfie.cluster=as.factor(cellfie.cluster))

cluster.cellfie.type <- as.data.frame(kmeans.cellfie.type$cluster) %>% 
  tibble::rownames_to_column(var="Sample_ID") %>%
  dplyr::rename("cellfie.type"="kmeans.cellfie.type$cluster") %>%
  mutate(cellfie.type=as.factor(cellfie.type))

pheno.clusters <- left_join(pheno, cluster.TPM) %>% left_join(cluster.cellfie) %>%
  left_join(cluster.TPM.type) %>% left_join(cluster.cellfie.type) %>%
  arrange(Cell_Type) 
```

bar plot cluster by immune cell
```{r}
# TPM
ggplot(data = pheno.clusters, aes(y = TPM.cluster)) +
  geom_bar(aes(fill = Immune_Cell)) +
  ggtitle("Clusters by Immune Cell Type using TPM") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(text=element_text(size=15))

# cellfie
ggplot(data = pheno.clusters, aes(y = cellfie.cluster)) +
  geom_bar(aes(fill = Immune_Cell)) +
  ggtitle("Clusters by Immune Cell Type using CellFie") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(text=element_text(size=15))
```
 
 bar plot immune cell by cluster
```{r}
# TPM
ggplot(data = pheno.clusters, aes(y =Immune_Cell)) +
  geom_bar(aes(fill = TPM.cluster)) +
  ggtitle("Clusters by Immune Cell Type using TPM") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(text=element_text(size=15))

# cellfie
ggplot(data = pheno.clusters, aes(y =Immune_Cell)) +
  geom_bar(aes(fill = cellfie.cluster)) +
  ggtitle("Clusters by Immune Cell Type using CellFie") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(text=element_text(size=15))
```

Clustering on subset of cells
```{r}
bob <- filter(pheno, Cell_Type == "Granulocytes" | Cell_Type == "Dentritic Cells" | Cell_Type == "NK-cells")
sam <- cellfie.matrix[, match(bob$unique_Imm, colnames(cellfie.matrix))]
identical(bob$unique_Imm, colnames(sam))

#remove rows/tasks with no variation
sam <- sam[matrixStats::rowSds(sam) > 0.000000001 ,]

set.seed(182)
kmeans.subset <- kmeans(scale(t(sam)), length(unique(bob$`Immune cell`)), nstart = 25)

cluster.subset <- as.data.frame(kmeans.subset$cluster) %>% 
  tibble::rownames_to_column(var="unique_Imm") %>%
  dplyr::rename("cluster"="kmeans.subset$cluster") %>%
  mutate(cluster=as.factor(cluster)) %>% right_join(bob)

ggplot(data = cluster.subset, aes(y = cluster)) +
  geom_bar(aes(fill = `Immune cell`)) +
  ggtitle("Clusters by Immune Cell Type using CellFie") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(text=element_text(size=15))
```




