---
title: "R Notebook"
output: html_notebook
---

```{r}
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE)
```

Load the required library packages
```{r}
library(data.table)
library(dplyr)
library(tidyr)
library(AnnotationDbi)
library(org.Hs.eg.db)
```

######################
TISSUE
######################

Load HPA tissue data
You can download the data from: https://www.proteinatlas.org/about/download
```{r}
tissue.data <- fread(paste('unzip -p', "Input/rna_tissue_hpa.tsv.zip"))
tissue.pheno <- fread(paste('unzip -p', "Input/rna_tissue_hpa_description.tsv.zip"))
```

Generate expression matrix compatible with CellFie
```{r}
# Isolate expression data and reshape the data from long format to wide format (rows=samples, columns=genes)
TPM.matrix <- tissue.data[, c("Gene", "Tissue", "nTPM")] %>%
  spread(Tissue, nTPM)

# Convert ensemble gene IDs to entrez IDs
gene.map <- AnnotationDbi::select(org.Hs.eg.db, keys=TPM.matrix$Gene, columns=c("ENTREZID"), keytype="ENSEMBL") %>% na.omit()
TPM.matrix.entrez <- left_join(TPM.matrix, gene.map, by=c("Gene"="ENSEMBL")) %>%
  dplyr::select(ENTREZID, everything()) %>% dplyr::select(-Gene) %>% na.omit()
  
# Aggregate duplicate genes by max value
TPM.entrez.max <- aggregate(TPM.matrix.entrez[,-1], by=list(TPM.matrix.entrez$ENTREZID), FUN=max) %>%
  dplyr::rename(entrezID = Group.1) %>% sapply(as.numeric) %>% as.data.frame()

# Add pseudo count
TPM.entrez.pseudo <- cbind(ENTREZID=TPM.entrez.max$entrezID, TPM.entrez.max[, 2:ncol(TPM.entrez.max)] + 0.00001)
```

Generate properties data matrix (phenotypic information for the samples)
```{r}
setdiff(names(TPM.entrez.pseudo), tissue.pheno$Tissue)

# remove " 1" from pheno Tissue names
tissue.pheno[tissue.pheno=="stomach 1"] <- "stomach"
tissue.pheno[tissue.pheno=="endometrium 1"] <- "endometrium"
tissue.pheno[tissue.pheno=="skin 1"] <- "skin"

setdiff(names(TPM.entrez.pseudo), tissue.pheno$Tissue)
setdiff(tissue.pheno$Tissue, names(TPM.entrez.pseudo))

# remove pheno data for samples not present in TPM matrix
tissue.pheno <- tissue.pheno[tissue.pheno$Tissue %in% names(TPM.entrez.pseudo)]
setdiff(tissue.pheno$Tissue, names(TPM.entrez.pseudo))
```

Save/Export files
```{r}
fwrite(TPM.entrez.pseudo, "Output/HPAtissue_expression.csv") #Archive: TPM.entrez.max
fwrite(tissue.pheno, "Output/HPAtissue_pheno.csv")
```


######################
BLOOD
######################

Load HPA immune cell data
```{r}
blood.data <- fread(paste('unzip -p', "Input/rna_immune_cell_sample.tsv.zip")) %>%
    # remove total PBMC
  filter(`Immune cell` != "total PBMC") %>%
  # sample ID as number causes issues downstream in Matlab... add prefix or suffix to force non-numeric reading
  mutate(`Sample ID` = paste0("sample_", `Sample ID`))

```

Generate expression matrix compatible with CellFie
```{r}
expression.type <- c("TPM", "pTPM", "nTPM")

for(exp in expression.type) {
 
  # Isolate expression data and reshape the data from long format to wide format (rows=samples, columns=genes)
  TPM.matrix <- dplyr::select(blood.data, c("ENSG ID", "Sample ID", exp)) %>%
    tidyr::spread_("Sample ID", exp)
  
  # Convert ensemble gene IDs to entrez IDs
  gene.map <- AnnotationDbi::select(org.Hs.eg.db, keys=TPM.matrix$`ENSG ID`, columns=c("ENTREZID"), keytype="ENSEMBL") %>%
    na.omit()
  TPM.matrix.entrez <- left_join(TPM.matrix, gene.map, by=c("ENSG ID"="ENSEMBL")) %>%
    dplyr::select(ENTREZID, everything()) %>% dplyr::select(-`ENSG ID`) %>% na.omit()
  
  # Aggregate duplicate genes by max value
  TPM.entrez.max <- aggregate(TPM.matrix.entrez[,-1], by=list(TPM.matrix.entrez$ENTREZID), FUN=max) %>%
  dplyr::rename(entrezID = Group.1) %>% sapply(as.numeric) %>% as.data.frame()
  
  # Add pseudo count
  TPM.entrez.pseudo <- cbind(ENTREZID=TPM.entrez.max$entrezID, TPM.entrez.max[, 2:ncol(TPM.entrez.max)] + 0.00001)
  
  # Export file
  fwrite(TPM.entrez.pseudo, paste0("Output/HPAblood_109_", exp, "expression.csv"))
}
```

Generate properties data matrix (phenotypic information for the samples)
```{r}
blood.pheno <- dplyr::select(blood.data, c("Sample ID", "Donor", "Immune cell")) %>% unique()
fwrite(blood.pheno, "Output/HPAblood_109_pheno.csv")
```
