---
title: "R Notebook"
output: html_notebook
---

```{r}
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE)
```

```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(treemapify)
library(RColorBrewer)
library(factoextra)
library(ReactomePA)
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)
library(ggraph)
library(tidygraph)
```

Load Curtis DESeq results
```{r}
DE <-fread("Input/DE.csv")
DEsig_0.05 <- DE[padj < 0.05 & abs(log2FoldChange) > log2(1.5)]
```

Load TPM data
```{r}
TPM.raw <- feather::read_feather("Input/abund.feather") 
TPM.cellfie <- fread("Input/Archive/TPM.FetuinB_pre.csv") # filtered to only include genes used in CellFie analysis
```


###################################################### 
CELLFIE ANALYSIS
######################################################

Load Cellfie results
```{r}
res.cellfie <- fread("CellFie_output/transcriptome/local_mean/score.csv") %>% setNames(c("NAFLD_R1", "NAFLD_R2", "NAFLD_R3", "WT_R1", "WT_R2", "WT_R3"))
taskInfo <-  fread("CellFie_output/transcriptome/local_mean/taskInfo.csv") %>%
  mutate(Pathway = case_when(stringr::str_detect(taskInfos1, "sec_") ~ 'Secretory',
         !stringr::str_detect(taskInfos1, "sec_") ~ 'Metabolism')) %>%
  setNames(c("taskID", "Task", "System", "Subsystem", "Pathway"))
detail.scoring <- fread("CellFie_output/transcriptome/local_mean/detailScoring.csv")

# Met only
res.cellfie.met <- cbind(taskInfo, res.cellfie) %>%
  filter(Pathway=="Metabolic")
taskInfo.met <- filter(taskInfo, Pathway=="Metabolic")

# Sec only
res.cellfie.sec <- cbind(taskInfo, res.cellfie) %>%
  filter(Pathway=="Secretory")
taskInfo.sec <- filter(taskInfo, Pathway=="Secretory")

# Cell line labels and colors
cellLine.labels <- as.factor(c(rep("NAFLD",3), rep("WT", 3)))
cellLine_col <- brewer.pal(2,"Dark2")[as.numeric(cellLine.labels)]

```

Differentially active tasks
```{r}
welch.test <- matrixTests::row_t_welch(res.cellfie.sec[, c("NAFLD_R1", "NAFLD_R2", "NAFLD_R3")], res.cellfie.sec[, c("WT_R1", "WT_R2", "WT_R3")]) %>% data.table() %>% cbind(taskInfo.sec) %>% mutate(p.adjust := p.adjust(pvalue, method="BH")) %>% mutate(FC := mean.x/mean.y) %>% mutate(LFC := log2(FC))
welch.test <- welch.test[, c("mean.x", "mean.y", "mean.diff", "FC" ,"LFC", "p.adjust", "Task", "System", "Subsystem", "Pathway")]


sig.tasks<- welch.test[, c("mean.x", "mean.y", "mean.diff", "FC" ,"LFC", "p.adjust", "Task", "System", "Subsystem", "Pathway")] %>% 
  filter(p.adjust <= 0.1) %>% 
  filter(LFC <= -log2(1.25) | LFC >= log2(1.25))

```

Volcano plot
```{r}
#selectLab <- sig.tasks$Task

selectLab <- paste0("italic('", sig.tasks$Task, "')")

png("Figures/volcano_cellFieTranscriptome.png", width=750, height=800)
EnhancedVolcano::EnhancedVolcano(welch.test, 
                                 lab=welch.test$Task,
                                 x= "LFC", 
                                 y="p.adjust", 
                                 axisLabSize = 30,
                                 pCutoff=0.1, 
                                 FCcutoff = log2(1.25),
                                 labSize = 8,
                                 title = "Secretory Pathway CellFie Volano Plot",
                                 titleLabSize = 30,
                                 subtitle = bquote(italic('NAFLD vs. WT')),
                                 subtitleLabSize = 28,
                                 ylim=c(0,5),
                                 xlim = c(-1, 1),
                                 pointSize = c(ifelse(welch.test$p.adjust<=0.1 & abs(welch.test$LFC) >= log2(1.25), 10, 5)),
                                 caption = bquote("fold change cutoff: 1.25; FDR cutoff: 0.1"),
                                 captionLabSize = 25,
                                 legendPosition = "bottom",
                                 legendLabSize = 28,
                                 colAlpha = 0.9,
                                 drawConnectors = T,
                                 widthConnectors = 0,
                                 colConnectors = rgb(0, 0, 0, max = 1, alpha = 0, names = ""),
                                 typeConnectors = "open",
                                 ylab = expression("-Log"[10]*" FDR"),
                                 legendLabels = c("NS", expression(Log[2] ~ FC), "FDR", expression(FDR ~ and ~ log[2] ~ FC))
                                 ) 
dev.off()
```

###################################################### 
PCA
######################################################

PCA using CellFie
```{r}
### prepare expression data for PCA
cellfie.matrix <- dplyr::select(res.cellfie.sec, c("NAFLD_R1", "NAFLD_R2", "NAFLD_R3", "WT_R1", "WT_R2", "WT_R3")) %>% as.matrix() %>% `rownames<-`(taskInfo.sec$Task)
# remove constant/zero columns
cellfie.matrix <- cellfie.matrix[matrixStats::rowSds(cellfie.matrix) != 0 ,]

# PCA
res.pca <- prcomp(t(cellfie.matrix), scale=T)

# visualize eigenvalues/variances
fviz_screeplot(res.pca, addlabels=T, ylim = c(0, 100))

# plot individuals on first two components
fviz_pca_biplot(res.pca,
             geom.ind = c("text", "point"),
             select.var = list(contrib = 4),
             col.var="gray59",
             #jitter = list(what = "label", height=0.2),
             arrowsize = 1,
             labelsize = 6,
             label="var",
             pointsize = 5,
             mean.point = FALSE,
             show.legend=FALSE,
             habillage = as.factor(cellLine.labels),
             palette = "Dark2",
             legend.title = "Condition",
             repel = TRUE     # Avoid text overlapping
) + theme(text=element_text(size=10), axis.text = element_text(size = 18), axis.title =
                           element_text(size = 18), legend.text = element_text(size = 18), legend.title =
                           element_text(size = 18), plot.title = element_text(size = 22)) +
  # ADD ggforce's ellipses
 ggforce::geom_mark_ellipse(aes(fill = cellLine.labels, color=cellLine.labels)) + #color = cellLine_col
  labs(title="Principal Component Analysis") + ylim(-4,4)

```


###################################################### 
Compare interactome with significant CellFie Tasks
######################################################

Load interacomics data
```{r}
## Raw LFQ data from Caressa
protLFQ <- readxl::read_excel("Input/Summary_Results_HEP.xlsx", sheet="Perseus")
protLFQ_trim <- data.table(protLFQ)[, c("Gene names","Welch's T-test Difference LFQ intensity TRAb_LFQ intensity WTAb", "Welch's T-test q-value LFQ intensity TRAb_LFQ intensity WTAb")] %>% setNames(c("geneName", "LFQ_FetB.vs.WT", "qvalue"))
#note: can't include protein names because number of proteins names does NOT equal number of gene names for split
protLFQ_parse <- mutate(protLFQ_trim, id=as.character(1:nrow(protLFQ_trim))) %>% tidyr::separate_rows(geneName, sep=";") %>% data.table()
protLFQ_sig <- protLFQ_parse[qvalue <= 0.1 & abs(LFQ_FetB.vs.WT) >= log2(1.5) ,][order(LFQ_FetB.vs.WT)]

## Raw interactomics data used for CellFie
interactome_raw <- fread("Output/FetuinB_interactomics.csv") %>% 
  setNames(c("Entrez", "NAFLD_R1", "NAFLD_R2", "NAFLD_R3", "WT_R1", "WT_R2", "WT_R3"))

## gene map (entrez - gene name)
entrezMap <- AnnotationDbi::select(org.Hs.eg.db, keys=as.character(interactome_raw$Entrez), keytype = "ENTREZID", columns="SYMBOL")
```

Load essential reactions data
```{r}
essRxn <- openxlsx::read.xlsx("../../usingGeM/Generate_essentialRxns_manual/essentialRxn_final2022.xlsx", sheet="Human.tasks")
```

Function to extract genes from Task GPR rules
```{r}
# extract all genes from GPR rules
parseGenes <- function(Task.df) {
  parseGPR.list <- stringr::str_extract_all(Task.df$`Gene.Protein.Reaction.(GPR).association.(Entrez.Gene.IDs)`, "\\d+") %>% 
  `names<-`(Task.df$Reaction.abbreviation)
  genes <- unlist(parseGPR.list, use.names=F)  %>% unique() %>% as.integer()
  return(genes)
}
```

List of tasks
```{r}
cnet.list <- list()

## Vesicle secretion
VS <- filter(essRxn, Task=="Vesicle secretion") %>% parseGenes()
cnet.list[["Vesicle secretion"]] <- filter(interactome_raw, Entrez %in% intersect(VS, interactome_raw$Entrez)) %>%
  mutate(Entrez=as.character(Entrez)) %>%
  left_join(entrezMap, by=c("Entrez"="ENTREZID")) %>%
  left_join(protLFQ_parse, by=c("SYMBOL"="geneName"))

## N-glycan processing (ER)
NG <- filter(essRxn, Task=="N-glycan processing (ER)") %>% parseGenes()
cnet.list[["N-glycan processing (ER)"]] <- filter(interactome_raw, Entrez %in% intersect(NG, interactome_raw$Entrez)) %>% 
  mutate(Entrez=as.character(Entrez)) %>%
  left_join(entrezMap, by=c("Entrez"="ENTREZID")) %>%
  left_join(protLFQ_parse, by=c("SYMBOL"="geneName"))

## Calnexin/calreticulin cycle
CAL <- filter(essRxn, Task=="Calnexin/calreticulin cycle") %>% parseGenes()
cnet.list[["Calnexin/calreticulin cycle"]] <- filter(interactome_raw, Entrez %in% intersect(CAL, interactome_raw$Entrez)) %>%
  mutate(Entrez=as.character(Entrez)) %>%
  left_join(entrezMap, by=c("Entrez"="ENTREZID")) %>%
  left_join(protLFQ_parse, by=c("SYMBOL"="geneName"))

## Trafficking to endosomes
ENDO <- filter(essRxn, Task=="Trafficking to endosomes (Clathrin Vesicles)") %>% parseGenes()
cnet.list[["Trafficking to endosomes (Clathrin Vesicles)"]] <- filter(interactome_raw, Entrez %in% intersect(ENDO, interactome_raw$Entrez)) %>% 
  mutate(Entrez=as.character(Entrez)) %>%
  left_join(entrezMap, by=c("Entrez"="ENTREZID")) %>%
  left_join(protLFQ_parse, by=c("SYMBOL"="geneName"))

cnet.df <- rbindlist(cnet.list, idcol="Task")
cnet.genes <- filter(DE, symbol.sep %in% cnet.df$SYMBOL) %>% unique() %>%
  dplyr::select(c("symbol.sep", "log2FoldChange", "padj")) 
```


```{r}
cnet.edge <- data.frame(from=cnet.df$Task, to=cnet.df$SYMBOL, interaction_strength=cnet.df$LFQ_FetB.vs.WT)
cnet.node <- dplyr::select(sig.tasks, c("Task", "LFC", "p.adjust")) %>% 
  setNames(names(cnet.genes)) %>%
  mutate(size=rep(9, nrow(sig.tasks))) %>%
  mutate(font=rep("bold", nrow(sig.tasks))) %>%
  rbind(cnet.genes %>% mutate(size=rep(6, nrow(cnet.genes))) %>% mutate(font=rep("italic", nrow(cnet.genes))))

graph <- igraph::graph_from_data_frame(cnet.edge, directed=FALSE, vertices=cnet.node)
```

```{r}
png("Figures/sigTask_interactome_nodeGradient.png", width=750, height=600)
set.seed(18)
ggraph(graph, layout="fr") +
  geom_edge_link(aes(color=interaction_strength), width=1.5) +
  geom_node_point(aes(size=size, color=log2FoldChange)) +
  scale_size_continuous(range = c(4, 12)) +
  geom_node_text(aes(label=name, fontface=font, size=6.8), repel=T) +
  scale_edge_colour_gradient2(low="red", mid='white', high="blue", name="Edge LFQ") +
  scale_colour_gradient2(low="red", mid='white', high="blue", name="Node LFC") +
  theme_void() +
  theme(text=element_text(size=22)) +
  guides(size=FALSE)
dev.off()
```


###################################################### 
Heatmap of CellFie Scores
######################################################

```{r}
### prepare expression data
cellfie.matrix <- dplyr::select(res.cellfie, c("NAFLD_R1", "NAFLD_R2", "NAFLD_R3", "WT_R1", "WT_R2", "WT_R3")) %>% as.matrix() %>% `rownames<-`(taskInfo$Task)
# remove constant/zero columns
cellfie.matrix <- cellfie.matrix[matrixStats::rowSds(cellfie.matrix) != 0 ,]

```

```{r}
png("Figures/hmap_tasks.png", width=550, height=550)
hmap2 <- pheatmap::pheatmap(cellfie.matrix,
                   scale="row",
                   color=colorRampPalette((c("red", "white", "blue")))(50),
                   cluster_rows = T, 
                   cluster_cols = T,
                   show_rownames = F,
                   cutree_cols = 2,
                   cutree_rows = 2,
                   treeheight_row = 0,
                   annotation_row = dplyr::select(clust.samples, c("Cluster")) %>% `rownames<-`(clust.samples$Task),
                   annotation_colors = list(Cluster=c("darkorchid1", "darkorange") %>% setNames(c("A", "B"))),
                   main="Clustered heatmap of CellFie scores",
                   fontsize=15,
                   angle_col = 45,
                   cellwidth = 60)
dev.off()
```

hypergeometric enrichment per cluster
```{r}
set.seed(18)
enrich_clusterA <- clusterProfiler::enricher(gene = clust.samples$Task[clust.samples$Cluster=="A"],
                                             TERM2GENE = dplyr::select(taskInfo, c("Subsystem", "Task")) %>%
                                               rbind(dplyr::select(taskInfo, c("System", "Task")), use.names=F),
                                             universe = clust.samples$Task,
                                             #minGSSize = 3,
                                             pAdjustMethod = "BH",
                                             pvalueCutoff = 0.05,
                                             qvalueCutoff = 0.1)
View(data.table(enrich_clusterA@result)[p.adjust <= 0.1 ,])

set.seed(18)
enrich_clusterB <- clusterProfiler::enricher(gene = clust.samples$Task[clust.samples$Cluster=="B"],
                                             TERM2GENE = dplyr::select(taskInfo, c("Subsystem", "Task")) %>%
                                               rbind(dplyr::select(taskInfo, c("System", "Task")), use.names=F),
                                             universe = clust.samples$Task,
                                             #minGSSize = 3,
                                             pAdjustMethod = "BH",
                                             pvalueCutoff = 0.05,
                                             qvalueCutoff = 0.1)
View(data.table(enrich_clusterB@result)[p.adjust <= 0.1 ,])
```

```{r}
enrich.res <- (data.table(enrich_clusterA@result)[p.adjust <= 0.1 ,] %>% mutate(Cluster="A")) %>% arrange(-Count) %>%
  rbind(data.table(enrich_clusterB@result)[p.adjust <= 0.1 ,] %>% mutate(Cluster="B")) %>% 
  mutate(FDR=paste("FDR =", formatC(p.adjust, format="e", digits=2))) %>% 
  mutate(term_count=sub("/.*","" , BgRatio)) %>%
  mutate(percent=(Count/as.integer(term_count))*100)
enrich.res$ID <- factor(enrich.res$ID, levels=rev(enrich.res$ID))

png("Figures/bar_enrichTasks.png", width=680, height=400)
ggplot(enrich.res, aes(x=ID, y=percent, fill=Cluster)) +
  geom_bar(stat="identity") +
  ylab("% tasks") +
  geom_text(aes(label = FDR), vjust = 0, hjust=1.1, size=4.5) +
  coord_flip() +
  labs(title="Hypergeometric Enrichment of Clusters") +
  scale_fill_manual(values=c("darkorchid1", "darkorange")) +
  theme_classic() +
  theme(text=element_text(size=17), axis.title.y = element_blank(), legend.position = "none",
        plot.title = element_text(size = 20, face = "bold"))
  #facet_wrap(~Cluster, scales="free")
dev.off()
```

