---
title: "R Notebook"
output: html_notebook
---

Clear packages
```{r}
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE)
```

Load packages
```{r}
library(data.table)
library(dplyr)
library(factoextra)
library(ggplot2)
library(gplots)
library(pheatmap)
library(RColorBrewer)
library(colorspace)
library(dendextend)
library(Rmisc)
```

Load original TPM data (for labels) and pheno data
```{r}
TPM <- fread("Output/HPAblood_109_TPMexpression.csv")
pheno <- fread("Output/HPAblood_109_pheno_withType.csv", colClasses=c("character"))

identical(colnames(TPM)[-1], pheno$Sample_ID)
pheno <- pheno[match(colnames(TPM)[-1], pheno$Sample_ID) ,]
identical(colnames(TPM)[-1], pheno$Sample_ID)
```

Load Cellfie results
```{r}
res.cellfie <- fread("CellFie_output/blood/metabolism_secretion/minmaxmean/109_pTPM/score.csv") %>%
  setNames(colnames(TPM)[-1])
res.cellfie.binary <- fread("CellFie_output/blood/metabolism_secretion/minmaxmean/109_pTPM/score_binary.csv") %>%
  setNames(colnames(TPM)[-1])
taskInfo <- fread("CellFie_output/blood/metabolism_secretion/minmaxmean/109_pTPM/taskInfo.csv") %>%
  mutate(Pathway = case_when(stringr::str_detect(taskInfos1, "sec_") ~ 'Secretory',
         !stringr::str_detect(taskInfos1, "sec_") ~ 'Metabolism')) %>%
  setNames(c("taskID", "Task", "System", "Subsystem", "Pathway"))

res.cellfie.filter <- res.cellfie * res.cellfie.binary
```

TPM matrix
```{r}
### prepare expression data
TPM.matrix <- as.matrix(TPM[, -1]) %>% `rownames<-`(TPM$ENTREZID) %>% `colnames<-`(pheno$Sample_ID)
# remove constant/zero columns
TPM.matrix <- TPM.matrix[matrixStats::rowSds(TPM.matrix) > 0.001 ,]

```

CellFie matrix
```{r}
### prepare expression data
cellfie.matrix <- as.matrix(res.cellfie) %>% `rownames<-`(taskInfo$Task) %>% `colnames<-`(pheno$Sample_ID)

# remove constant/zero columns
cellfie.matrix <- cellfie.matrix[matrixStats::rowSds(cellfie.matrix) != 0 ,]

```

####################################################################
Hierarchical Clustering
####################################################################

It is very difficult to get 18 distinct colors
Manually attempt to get most distinct colors as possible
```{r}
library(randomcoloR)
palette <- distinctColorPalette(18)
#pie(rep(1, 18), col=palette)

ann_colors <- list(cell=c(basophil="#E3589A", `classical monocyte`="#DA4FDF", eosinophil="#D86049", `gdT-cell`="#764ADF", `intermediate monocyte`="#7AE94E", `MAIT T-cell`="#D3BEDC", `memory B-cell`="#7FBFDC", `memory CD4 T-cell`="#D291E2", `memory CD8 T-cell`="#7E5C99", `myeloid DC`="#E0989B", `naive B-cell`="#C4E090", `naive CD4 T-cell`="#DEAF66",
`naive CD8 T-cell`="#D8DF52", neutrophil="#DDE0CC", `NK-cell`="#7DE0D1", `non-classical monocyte`="#809173", `plasmacytoid DC`="#6CDE90", `T-reg`="#6E95DE"))
```

Using all cellfie (met+sec)
```{r}

## no subtype annotations
cellType.col <- list(cellType=brewer.pal(length(unique(pheno$Cell_Type)), "Set2") %>%
  setNames(unique(pheno$Cell_Type)))

set.seed(18)
p.sec <- pheatmap::pheatmap(cellfie.matrix,
                    fontsize=6,
                    main="secCellFie", 
                    scale="row",
                    border_color = NA,
                    color=colorRampPalette(c("navy", "white", "red"))(50),
                    clustering_distance_cols = "correlation",
                    clustering_method = "complete",
                    annotation_col = data.frame(cellType=pheno$Cell_Type) %>% `rownames<-`(pheno$Sample_ID),
                    annotation_colors = cellType.col,
                    show_colnames = F,
                    show_rownames = F,
                    treeheight_row = 0,
                    legend_breaks = c(0,1))

## with subtype annotations
set.seed(18)
p.sec2 <- pheatmap::pheatmap(cellfie.matrix,
                    fontsize=6,
                    main="secCellFie", 
                    scale="row",
                    border_color = NA,
                    color=colorRampPalette(c("navy", "white", "red"))(50),
                    clustering_distance_cols = "correlation",
                    clustering_method = "complete",
                    annotation_col = data.frame(cell=factor(pheno$Immune_Cell), cellType=pheno$Cell_Type) %>% `rownames<-`(pheno$Sample_ID),
                    annotation_colors = append(ann_colors, cellType.col),
                    show_colnames = F,
                    show_rownames = F,
                    treeheight_row = 0,
                    legend_breaks = c(0,1))

## DC only subtype annotations
set.seed(18)
p.sec3 <- pheatmap::pheatmap(cellfie.matrix,
                    fontsize=6,
                    main="secCellFie", 
                    scale="row",
                    border_color = NA,
                    color=colorRampPalette(c("navy", "white", "red"))(50),
                    clustering_distance_cols = "correlation",
                    clustering_method = "complete",
                    annotation_col = data.frame(cell=factor(pheno$Immune_Cell), cellType=pheno$Cell_Type) %>% `rownames<-`(pheno$Sample_ID) %>% mutate(cell=case_when(cell=='plasmacytoid DC' ~ 'plasmacytoid DC',
                                                                                                                                                                      cell=='myeloid DC' ~ 'myeloid DC',
                                                                                                                                                                      TRUE ~ NA)),
                    annotation_colors = append(cellType.col, list(cell=c(`myeloid DC`="#E0989B",`plasmacytoid DC`="#6CDE90"))),
                    show_colnames = F,
                    show_rownames = F,
                    treeheight_row = 0,
                    legend=F,
                    annotation_legend = F)
```

Using met cellfie only
```{r}
met.taskInfo <- filter(taskInfo, Pathway=="Metabolism")

## no subtype annotations
p.met <- pheatmap::pheatmap(cellfie.matrix[rownames(cellfie.matrix) %in% met.taskInfo$Task ,],
                    fontsize=6,
                    main="CellFie", 
                    scale="row",
                    border_color = NA,
                    color=colorRampPalette(c("navy", "white", "red"))(50),
                    #cluster.cols = hclust(dist(cellfie.matrix, method="euclidean"), method="complete"),
                    clustering_distance_cols = "correlation",
                    clustering_method = "complete",
                    annotation_col = data.frame(cellType=pheno$Cell_Type) %>% `rownames<-`(pheno$Sample_ID),
                    annotation_colors = cellType.col,
                    show_colnames = F,
                    show_rownames = F,
                    treeheight_row = 0,
                    legend=F,
                    annotation_legend = F)

## with subtype annotations
set.seed(18)
p.met2 <- pheatmap::pheatmap(cellfie.matrix[rownames(cellfie.matrix) %in% met.taskInfo$Task ,],
                    fontsize=6,
                    main="CellFie", 
                    scale="row",
                    border_color = NA,
                    color=colorRampPalette(c("navy", "white", "red"))(50),
                    clustering_distance_cols = "correlation",
                    clustering_method = "complete",
                    annotation_col = data.frame(cell=factor(pheno$Immune_Cell), cellType=pheno$Cell_Type) %>% `rownames<-`(pheno$Sample_ID),
                    annotation_colors = append(ann_colors, cellType.col),
                    show_colnames = F,
                    show_rownames = F,
                    treeheight_row = 0,
                    legend_breaks = c(0,1))

## DC only subtype annotations
set.seed(18)
p.met3 <- pheatmap::pheatmap(cellfie.matrix[rownames(cellfie.matrix) %in% met.taskInfo$Task ,],
                    fontsize=6,
                    main="CellFie", 
                    scale="row",
                    border_color = NA,
                    color=colorRampPalette(c("navy", "white", "red"))(50),
                    clustering_distance_cols = "correlation",
                    clustering_method = "complete",
                    annotation_col = data.frame(cell=factor(pheno$Immune_Cell), cellType=pheno$Cell_Type) %>% `rownames<-`(pheno$Sample_ID) %>% mutate(cell=case_when(cell=='plasmacytoid DC' ~ 'plasmacytoid DC',
                                                                                                                                                                      cell=='myeloid DC' ~ 'myeloid DC',
                                                                                                                                                                      TRUE ~ NA)),
                    annotation_colors = append(cellType.col, list(cell=c(`myeloid DC`="#E0989B",`plasmacytoid DC`="#6CDE90"))),
                    show_colnames = F,
                    show_rownames = F,
                    treeheight_row = 0)

```

need to use grid.draw function with pheatmaps
function to save pheatmap images
```{r}
save_pheatmap_png <- function(x, filename, width, height, res) {
  png(filename, width = width, height = height, res = res)
  grid::grid.newpage()
  grid::grid.draw(x$gtable)
  dev.off()
}
```

```{r}
# plot with no legends to get equal size dendrogram
save_pheatmap_png(p.sec3, "Figures/hclust_CellFie_sec3.png", width=1000, height=800, res=350)
save_pheatmap_png(p.met, "Figures/hclust_CellFie_met.png", width=1000, height=800, res=350)

# plot with legend
save_pheatmap_png(p.met3, "Figures/hclust_CellFie_met3.png", width=1200, height=800, res=300)

```

Compare dendrograms
```{r} 
### dendrogram matching known lineage with manually generated distance matrix
HPAdist.df <- openxlsx::read.xlsx("Output/HPAblood_109_distMatrix.xlsx", colNames=F)
HPA.dist <- as.matrix(HPAdist.df[-c(1:4), -c(1:4)]) %>%
  apply(2, as.numeric) %>% 
  Matrix::forceSymmetric(uplo="L") %>%
  as.matrix() %>%
  `colnames<-`(HPAdist.df$X1[-c(1:4)]) %>%
  `rownames<-`(HPAdist.df$X1[-c(1:4)])
hclust.HPA <- hclust(as.dist(HPA.dist))
#plot(hclust.HPA)

```

tanglegram
```{r}
dend.met <- as.dendrogram(p.met$tree_col)
dend.sec <- as.dendrogram(p.sec$tree_col)
dend.HPA <- as.dendrogram(hclust.HPA)

dendList.met <- dendlist(dend.met, dend.HPA)
dendList.sec <- dendlist(dend.sec, dend.HPA)

set.seed(18)
dendList.met %>% #untangle(method="random", R=1) %>% 
  untangle(method="DendSer") %>%
  tanglegram(sort=F,
             highlight_distinct_edges = FALSE, # Turn-off dashed lines
             common_subtrees_color_lines = FALSE, # Turn-off line colors
             common_subtrees_color_branches = FALSE # Color common branches
             ) %>% entanglement()
set.seed(18)
dendList.sec %>% #untangle(method="random", R=1) %>% 
  untangle(method="DendSer") %>%
  tanglegram(sort=F,
             highlight_distinct_edges = FALSE, # Turn-off dashed lines
             common_subtrees_color_lines = FALSE, # Turn-off line colors
             common_subtrees_color_branches = FALSE # Color common branches
             ) %>% entanglement()

```

correlations
```{r}
cor.dendlist(dendList.met, method="cophenetic")
cor.dendlist(dendList.met, method="baker")

cor.dendlist(dendList.sec, method="cophenetic")
cor.dendlist(dendList.sec, method="baker")
```

```{r}
dendList <- dendlist(`CellFie`=dend.met, `secCellFie`=dend.sec, HPA=dend.HPA)
cor.dend <- cor.dendlist(dendList, method="FM_index", k=6)
round(cor.dend, 2)

png("Figures/dend_corr.png", width=600, height=400)
corrplot::corrplot(cor.dend, method="circle", addCoef.col = 'white', col=corrplot::COL1('YlOrBr', 100), type='upper', tl.col = "black", tl.cex=1.5, cl.cex=1.5, cl.align.text = "l", number.cex=1.5, tl.srt=45)
dev.off()

```



You can clearly see that certain tasks delineare the two main cluster of cells
Interestingly the two subtypes of DC cells fall in different clusters
Cluster tasks to assess how these two main clusters of cells are different (especially pDC vs mDC)

#####
Follow up (DENDRITIC only)
Isolate DC samples and perform clustering on tasks

```{r}
nclust <- 2
dendritic <- cellfie.matrix[, pheno$Sample_ID[pheno$Cell_Type=="Dendritic Cells"]]
dendritic <- dendritic[matrixStats::rowSds(dendritic) > 0.0001 ,]

# hierarchcial clustering object that will be used to annotate clusters of plot used for publication
set.seed(18)
hclust.dendritic <- pheatmap::pheatmap(dendritic,
                    fontsize=8, #20
                    main="Hierarchical clustering using secretory expansion of CellFie", 
                    scale="row",
                    border_color = NA,
                    color=colorRampPalette(c("navy", "white", "red"))(50),
                    #cluster.cols = hclust(dist(cellfie.matrix, method="euclidean"), method="complete"),
                    clustering_distance_cols = "correlation",
                    clustering_method = "complete",
                    annotation_col = data.frame(cell=factor(pheno$Immune_Cell), cellType=pheno$Cell_Type) %>% `rownames<-`(pheno$Sample_ID),
                    #annotation_row = data.frame( cluster=cutree(hclust.cellfie_clust$tree_row, k=nclust)) %>%
                      #`rownames<-`(rownames(cellfie.matrix)),
                    #annotation_colors = ann_colors,
                    show_colnames = F,
                    show_rownames = F,
                    #treeheight_row = 0,
                    cutree_rows=nclust)
```

use information from hclust.dendritic to add color annotations to plot for publication:
```{r}
clust.dendritic <- data.frame(clusterID=cutree(hclust.dendritic$tree_row, k=nclust), 
                                   Task=rownames(dendritic)) %>%
  `rownames<-`(rownames(dendritic)) %>% mutate(cluster=paste0("cluster",clusterID))
clust.dendritic.sort <- clust.dendritic[hclust.dendritic$tree_row$order ,] %>% left_join(taskInfo)

ann_colors <- list(cluster=c(cluster1="goldenrod2", cluster2="lightblue3"),
                               cell=c(`plasmacytoid DC`="#6CDE90", `myeloid DC`="#E0989B"))

set.seed(18)
hclust.dendritic.clust <- pheatmap::pheatmap(dendritic,
                    fontsize=20, #20
                    main="Dendritic Cells", 
                    scale="row",
                    border_color = NA,
                    color=colorRampPalette(c("navy", "white", "red"))(50),
                    #cluster.cols = hclust(dist(cellfie.matrix, method="euclidean"), method="complete"),
                    clustering_distance_cols = "correlation",
                    clustering_method = "complete",
                    annotation_col = data.frame(cell=factor(pheno$Immune_Cell)) %>% `rownames<-`(pheno$Sample_ID),
                    annotation_row = select(clust.dendritic, cluster),
                    annotation_colors = ann_colors,
                    show_colnames = F,
                    show_rownames = F,
                    #treeheight_row = 0,
                    cutree_rows=nclust,
                    treeheight_row=0)

save_pheatmap_png(hclust.dendritic.clust, "Figures/hclust_DC.png", width=700, height=900, res=80)
```

Plot cluster trends
```{r}
task_cluster <- dplyr::select(clust.dendritic, c("cluster", "Task")) %>%
  inner_join(data.frame(Task=rownames(dendritic), t(scale(t(dendritic))))) #%>% setcolorder(c("gene", "cluster", rownames(scale.data)[hmap$colInd]))
#convert wide to long
task_cluster.long <- tidyr::gather(task_cluster, sample, score, sample_1004:sample_992, factor_key=F)

### Too noisy to plot every task. Summarize trend with mean.
task_cluster.summary <- plyr::ddply(task_cluster.long, c("cluster", "sample"), summarise,
                              mean=CI(score, ci=0.95)['mean'],
                              sd=sd(score),
                              se=plotrix::std.error(score),
                              uci=CI(score, ci=0.95)['upper'],
                              lci=CI(score, ci=0.95)['lower'])
cluster.1 <- task_cluster.summary[task_cluster.summary$cluster=="cluster1" ,] %>%
  arrange(match(sample, hclust.dendritic.clust$tree_col$labels[hclust.dendritic.clust$tree_col$order]))
cluster.2 <- task_cluster.summary[task_cluster.summary$cluster=="cluster2" ,] %>%
  arrange(match(sample, hclust.dendritic.clust$tree_col$labels[hclust.dendritic.clust$tree_col$order]))

```

Plot boxplot of cluster trends
```{r}
box.cluster <- left_join(task_cluster.long,
                          select(pheno, c("Sample_ID", "Immune_Cell")), by=c("sample"="Sample_ID"))

png("Figures/DCcluster_boxplots.png", width = 800, height = 350, res=150)
ggplot(box.cluster, aes(x=Immune_Cell, y=score, fill=Immune_Cell)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#E0989B", "#6CDE90")) +
  theme(axis.ticks.x = element_blank()) +
  facet_wrap(~cluster, scale="free")
dev.off()
```


Tasks in dendritic clusters
```{r}
dendritic.clusters <- clust.dendritic.sort %>% left_join(taskInfo)
openxlsx::write.xlsx(dendritic.clusters, "Output/DC_clusters.xlsx")
#fwrite(filter(dendritic.clusters, Pathway=="Metabolism"), "Output/DC_Metabolism.csv")
```

Hypergeometric enrichment of cellfie systems in each cluster
```{r}
set.seed(18)
enrich.1 <- clusterProfiler::enricher(clust.dendritic.sort$Task[clust.dendritic.sort$clusterID==1], 
                                      TERM2GENE=dplyr::select(taskInfo, c("Pathway", "Task")),
                                      universe=clust.dendritic.sort$Task,
                                      minGSSize=3,
                                      #maxGSSize=16,
                                      pAdjustMethod = "fdr",
                                      qvalueCutoff = 0.02)

set.seed(18)
enrich.2 <- clusterProfiler::enricher(clust.dendritic.sort$Task[clust.dendritic.sort$clusterID==2],
                                      TERM2GENE=dplyr::select(taskInfo, c("Pathway", "Task")),
                                      universe=clust.dendritic.sort$Task,
                                      minGSSize=3,
                                      #maxGSSize=16,
                                      pAdjustMethod = "fdr",
                                      qvalueCutoff = 0.2)

#enrich_cor.Stress.sig <- enrich_cor.Stress@result %>% subset(p.adjust <= 0.1)
#rm(enrich_cor.Stress)
View(as.data.frame(enrich.1@result)%>% subset(p.adjust <= 0.1))
View(as.data.frame(enrich.2@result)%>% subset(p.adjust <= 0.1))
```

```{r}
clust.1 <- filter(clust.dendritic.sort, clusterID==1)
clust.2 <- filter(clust.dendritic.sort, clusterID==2)

clust.percent.table <- table(clust.1$System) %>% as.data.frame() %>% `colnames<-`(c("System", "Cluster_1")) %>%
  full_join(table(clust.2$System) %>% as.data.frame() %>% `colnames<-`(c("System", "Cluster_2"))) %>%
  full_join(table(clust.dendritic.sort$System) %>% as.data.frame() %>% `colnames<-`(c("System", "Total"))) %>%
  tidyr::replace_na(list(Cluster_1 = 0, Cluster_2 = 0)) %>%
  left_join(select(taskInfo, c("System", "Pathway")) %>% unique()) %>%
  arrange(Pathway)
             
```

Differential task activity
```{r}
welch.tasks <- function(cellfie.score, group1, group2, alt) {
  cellfie.score <- log2(cellfie.score+ 0.00001)
  group1.score <- cellfie.score[names(cellfie.score) %in% group1]
  group2.score <- cellfie.score[names(cellfie.score) %in% group2]
  welch.test <- t.test(group1.score, group2.score, alternative=alt)
  welch.res <- data.frame(p.value=welch.test$p.value, x.mean=welch.test$estimate[1], y.mean=welch.test$estimate[2])
  return(welch.res)
}
```

```{r}
DE_DC <- apply(dendritic, 1, welch.tasks, 
                        group1=pheno$Sample_ID[pheno$Immune_Cell=="myeloid DC"],
                        group2=pheno$Sample_ID[pheno$Immune_Cell=="plasmacytoid DC"],
                        alt="two.sided") %>%
  rbindlist(idcol = "Task") %>% dplyr::rename(myeloid.mean=x.mean) %>% dplyr::rename(plasmacytoid.mean=y.mean) %>% mutate(LFC=(myeloid.mean-plasmacytoid.mean)) %>%
  mutate(FDR=p.adjust(p.value, method="BH")) %>%
  left_join(taskInfo) %>%
  arrange(-LFC) 

DE_DC.sig <- DE_DC[FDR <= 0.1 & abs(LFC) >= log2(1.25)]
```

Volcano plot
```{r}
keyvals.colour <- case_when(
  DE_DC$Task %in% DE_DC.sig$Task[DE_DC.sig$Pathway=="Secretory"] ~ "#008B45",
  DE_DC$Task %in% DE_DC.sig$Task[DE_DC.sig$Pathway=="Metabolism"] ~ "#8B008B",
      .default="black")
names(keyvals.colour)[keyvals.colour == "#008B45"] <- 'Secretory Tasks'
names(keyvals.colour)[keyvals.colour == "#8B008B"] <- 'Metabolic Tasks'
names(keyvals.colour)[keyvals.colour == 'black'] <- 'Non-signficant'

png("Figures/volcano_HPA.png", width=1000, height=600)
EnhancedVolcano::EnhancedVolcano(DE_DC, 
                                 lab=DE_DC$Task, 
                                 colCustom = keyvals.colour,
                                 x= "LFC", 
                                 y="FDR", 
                                 axisLabSize = 24,
                                 pCutoff=0.1, 
                                 FCcutoff = log2(1.25),
                                 labSize = 0,
                                 title = "Dendritic Cell Volano Plot",
                                 titleLabSize = 30,
                                 subtitle = bquote(italic('myeloid DCs vs. plasmacytoid DCs')),
                                 subtitleLabSize = 24,
                                 ylim=c(0,5),
                                xlim=c(-8.5, 8.5),
                                pointSize = 4,
                                 caption = bquote(~Log[2]~ "fold change cutoff: 1.25; FDR cutoff: 0.1"),
                                 captionLabSize = 20,
                                 legendPosition = "right",
                                 legendLabels = c("NS", ~Log[2]~"FC", "FDR", ~Log[2]~"FC and FDR"),
                                 legendLabSize = 24,
                                 colAlpha = 0.9,
                                 ylab = expression("-Log"[10]*" FDR")
                                 ) #+ theme_classic()
dev.off()
```



